### 2020-05-28 Manuscript.Rmd
Init

---
title: "An Audio-Based 3D Spatial Guidance AR System for Blind Users"
titlerunning: Audio-Based 3D Spatial Guidance
authorrunning: J. Couhlan et al.
output:
  pdf_document:
    template: res/templates/markdown.tex
    toc: no
    keep_tex: yes
institutes:
- id: 1
  name: The Smith-Kettlewell Eye Research Institute, San Francisco, CA
  emails: coughlan@ski.org, brandon.biggs@ski.org, hshen@ski.org
- id: 2
  name: LITIS, University of Rouen, Normandy, France
  emails: marc-aurele.riviere@univ-rouen.fr
keywords:
- Assisitive Devices
- Augmented Reality
- Auditory Substitution
- Visual Impairment
abstract: Augmented reality (AR) has great potential for blind users because it enables
  a range of applications that provide audio information about specific locations
  or directions in the user’s environment. For instance, the CamIO (“Camera Input-Output”)
  AR app makes physical objects (such as documents, maps, devices and 3D models) accessible
  to blind and visually impaired persons by providing real-time audio feedback in
  response to the location on an object that the user is touching (using an inexpensive
  stylus). An important feature needed by blind users of AR apps such as CamIO is
  a 3D spatial guidance feature that provides real-time audio feedback to help the
  user find a desired location on an object. We have devised a simple audio interface
  to provide verbal guidance towards a target of interest in 3D. The experiment we
  report with blind participants using this guidance interface demonstrates the feasibility
  of the approach and its benefit for helping users find locations of interest.
authors:
- name: James Coughlan
  orcid: 0000-1111-2222-3333
  inst: 1
- name: Brandon Biggs
  orcid: 0000-1111-2222-3333
  inst: 1
- name: Marc-Aurèle Rivière
  orcid: 0000-0002-5108-3382
  inst: 2
- name: Huiying Shen
  orcid: 0000-1111-2222-3333
  inst: 1
ressources: res/
templates: res/templates/
bibliography: res/refs.bib
---

```{r setup, include=FALSE}

#TODO: DESCRIPTION with package metadata ?

knitr::opts_chunk$set(
  fig.show="hold",
  fig.align="default",
  warning = FALSE,
  echo = FALSE,
  cache = FALSE,
  error = FALSE
)

colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, x)
  } else x
}
```

# State of the Art and Related Technology {#sota}

Many people who are blind or visually impaired have difficulties accessing a range of everyday objects, including printed documents, maps, infographics, appliances and 3D models used in STEM education, needed for daily activities in schools, the home and the workplace. This limits their participation in many cultural, professional and educational activities. While braille labeling is often a useful means of providing access to such objects, there is only limited space for braille, and this method is only accessible to those who can read braille.

Audio labels are a powerful supplement or alternative to braille labels. Standard methods of implementing audio labels require the use of special hardware and materials, which is costly and limits their adoption. For instance, tactile graphics may be overlaid on a touch-sensitive tablet \cite{TalkingTactileTablet}. Documents or other surfaces can be covered with a special form of paper readable by a “smart pen” \cite{mieleTalkingTactileApps2010}. The PenFriend \cite{kendrickPenFriendTouchMemo2011} stylus uses small barcodes affixed to an object to define audio labels.

There is growing interest in using computer vision-based approaches for audio labeling. In these approaches a camera tracks the user’s fingers or handheld pointer as they explore an object, enabling audio labeling for existing objects with minimal or no customization. Past computer vision approaches to audio labeling include \cite{shiMarkitTalkitLowBarrier2017}, which focuses on 3D printed objects, \cite{thevinAugmentedRealityPeople2018}, which uses a depth camera to access flat documents, and \cite{fuscoTactileGraphicsHelper2015}, which provides access to tactile graphics to students in educational settings.

We build on our past work on CamIO \cite{shenCamIO3DComputer2013,coughlanEvaluatingAuthorUser2017,coughlanTowardsAccessibleAudioLabeling2020}, which is short for “Camera Input-Output.” We have implemented CamIO as a stand-alone iPhone app that tracks an inexpensive wooden stylus held by the user to point at locations of interest on an object (called hotspots). When the user points at a hotspot with the stylus tip, relevant audio information, such as text-to-speech or other sounds, is announced.

Experiments with blind participants have demonstrated the feasibility of using CamIO to access existing audio labels. Moreover, our approach allows a visually impaired user to create audio labels independently \cite{coughlanTowardsAccessibleAudioLabeling2020}. However, an ability that CamIO still lacks is the ability to provide real-time guidance to a hotspot. Without such guidance, the user must search the entire surface of an object with the stylus tip until they hear the hotspot they are seeking. This need is particularly acute in cases where the surface is lacking in tactile cues, such as the touch screen buttons on a microwave oven \cite{coughlanTowardsAccessibleAudioLabeling2020}.

Guidance not only saves the user time in exploring large or intricate objects, or objects lacking in tactile features, but can enable entirely new ways of interacting with objects. For instance, a geography learning app could ask a student to point to a specific city on a world globe, and if the student has trouble finding it, the app could guide them to it in an intuitive way. Other examples could include guiding a user to a specific button on an appliance (such as a microwave) to perform a desired function, guiding a student to a specific element on the periodic table, or even guiding a repairperson to find a part of a machine with limited visibility.
Our guidance system is inspired by recent work on auditory and haptic displays for 3D spatial guidance \cite{mayAuditoryDisplaysFacilitate2019,guezou-philippePrototypingEvaluatingSensory2018}, which provide real-time audio/haptic feedback to guide a user towards a target in a virtual environment. We initially developed an interface similar to \cite{mayAuditoryDisplaysFacilitate2019} for CamIO, but after preliminary testing we devised a simpler and more effective interface for our application, which we evaluated experimentally with four blind participants.

# Overview of the CamIO System {#overview}

CamIO is an augmented reality (AR) iPhone app that uses computer vision to provide audio labels for rigid 3D objects. The rigid object of interest is mounted on a board, which is a printed barcode pattern (\ref{fig:exp}) that allows CamIO to estimate the object’s pose (3D translation and 3D orientation) relative to the camera. The iPhone is typically mounted on a tripod or a gooseneck cellphone holder \ref{fig:exp} to obtain a clear view of the board and object, but in some cases \cite{coughlanTowardsAccessibleAudioLabeling2020} users may hold the iPhone by hand. The user points the tip of a passive wooden stylus \ref{fig:exp}, covered with another barcode pattern, to different locations on an object. CamIO estimates the stylus tip’s 3D location relative to the camera and uses the pose estimate of the board to determine the stylus tip’s 3D location relative to the object. Whenever the tip is close enough to a pre-defined hotspot, it triggers audio information about the hotspot. (In other words, hotspot locations are defined relative to the board, which is mounted rigidly to the object.)

A second stylus (not shown, identical in form but distinct from the one shown in \ref{fig:exp}) is used to create new hotspots, and allows the user to make an audio recording to associate with each new hotspot. Our recent work \cite{coughlanTowardsAccessibleAudioLabeling2020} demonstrates the accessibility of this authoring process, which allows blind users to create their own annotations (hotspots and associated audio labels) independently. However, in the study presented in this paper, the experimenter was the only one who created hotspots, since the emphasis here is on providing guidance to pre-existing hotspots.

# 3D Spatial Guidance Interface {#material}

We build on the auditory display described in \cite{mayAuditoryDisplaysFacilitate2019}, which is intended for audio-based 3D targeting of locations in a large virtual workspace (such as a kitchen). Accordingly, the first incarnation of our guidance approach used a repeating monaural sound, whose tempo indicated how far the stylus tip is from the target, and with a pitch (low, medium or high) that indicated whether the tip needed to move up or down (or else was aligned properly in the vertical direction).

Informal tests with two blind participants (one female of age 42, who is also participant P1 in the experiment described in the next section, and one male of age 73) showed that this guidance worked satisfactorily in some cases. However, we found that the feedback was too indirect: the participants didn’t always know how to move the stylus to increase the tempo. As a result, the resulting guidance process was unacceptably slow.
Thus we devised a second, more direct guidance approach: verbal directions telling the user to move the stylus along the cardinal directions (up, down, left, right, forward or back). While the optimal direction to move the stylus tip towards the hotspot might entail a combination of these directions (e.g., simultaneously move left and up), for simplicity we issue directions along only one cardinal direction at a time – the direction that will most improve the alignment to the target.

We define the guidance interface as follows. The board (barcode pattern beneath the truck, see \ref{fig:exp}) defines an xyz coordinate frame, with $+x$ = right, $-x$ = left, $+y$ = forward (away from the user), $-y$ = back (towards the user), $+z$ = up and $-z$ = down. Let the desired target hotspot location be denoted $(x*, y*, z*)$, and the current location of the stylus tip be $(x,y,z)$. Define the error vector $e = (x-x*, y-y*, z-z*)$, and let index $a = argmax_i |e_i|$, i.e., a is the index (1, 2 or 3, corresponding to x, y or z, respectively) of the entry in e with the highest absolute value. In other words, a indicates which coordinate (x, y or z) is most discrepant from the target coordinate location $(x*, y*, z*)$. Then the appropriate direction is issued, e.g., “left” if $a = 1$ and $x > x*$, “right” if $a = 1$ and $x < x*$, “up” if $a = 3$ and $z < z*$, etc. Directions are announced roughly twice per second whenever the stylus is visible. The target hotspot is announced when the stylus gets to within approximately 1 cm from the hotspot. No audio feedback is issued when the stylus is not visible.

# Experiment with Blind Participants {#exp}

We conducted an experiment to assess the feasibility of the spatial guidance feedback and to determine whether the feedback speeds up the target search. The experiment \ref{fig:exp} compared how much time a blind participant needed to find hotspots (points of interest) on a 3D object with spatial guidance vs. without it. The object was a large toy fire truck (27 cm x 12 cm x 14 cm), with K=10 distinct hotspot locations defined in advance. The hotspots were tactilely salient features such as lights, hooks or other features that protruded from the truck. Some hotspots were in recessed locations on the truck, such as the shift stick and steering wheel (\ref{fig:exp}) in the front cabin, or on the side of the truck facing away from the participant; these hotspots were difficult to find and access, whether by fingertip or by stylus. In fact, the difficulty of accessing such locations highlights an advantage of the CamIO stylus, whose tip need not be visible to the camera. This allows the user to explore recessed areas or areas that are otherwise invisible to the camera.

```{r exp, out.width=c("55%","45%"), fig.cap="(left) Blind participant uses CamIO to explore a toy fire truck. The participant holds the CamIO stylus (a barcode pattern printed on a square wooden dowel with a pointy tip) to point to possible hotspots on the truck. The truck is glued to the board, a piece of paper with a separate barcode pattern, which is taped to the table. An iPhone 8 running CamIO is held by a gooseneck cellphone holder to capture a view of the truck, board and stylus. (right) Stylus points to steering wheel in the truck cabin. Though the steering wheel is in a recessed location that is not visible to CamIO, the stylus tip location can still be estimated."}

knitr::include_graphics("res/img/fig1a.jpg")
knitr::include_graphics("res/img/fig1b.jpg")
```

Each hotspot location was searched for under two feedback conditions: G (guidance provided) and NG (no guidance). The participant was given up to 120 sec. to find each hotspot, after which a time-out was declared. In order to minimize the influence of any possible learning effects, which could make it easier to find a hotspot the second time it was searched for, we randomized the sequence of hotspots and alternated between NG and G trials. This resulted in a sequence of 2K = 20 trials, starting with an NG trial, the same sequence for each participant.

We had a total of N=4 blind participants in our experiment (2 male/2 female, ages from 26-42). Participants P1, P3 and P4 were already familiar with the CamIO system (P4 is a co-author of this paper) while P2 was unfamiliar with it. None of the participants was told the study design before their trials, and we didn’t tell them whether or not any locations would be repeated in the 20 trials. After obtaining IRB consent, we reviewed how CamIO works if necessary and demonstrated the audio guidance feature. The experimenter launched CamIO on an iPhone 8 and positioned it on a gooseneck cellphone holder so that the camera captured the entire truck and board.
Each participant spent a short time practicing with the guidance feature, with help provided by the experimenter as needed. (The experimenter chose practice hotspots that were different from the pre-defined K=10 hotspots.) In the practice phase, the experimenter explained that barriers sometimes block the user from moving in the suggested guidance direction. For instance, if in \ref{fig:exp} the system directs the participant to move “forward” towards the left side mirror, the stylus will soon hit the cabin roof/wall as it moves towards the mirror; the participant must find ways around such obstacles to find the target hotspot.

Next each participant performed the formal experimental trials, which consisted of the 2K=20 trial sequence described above. This sequence included 10 trials of both conditions (G and NG), for a total of 20 trials per participant. After completing the 20 search trials, we administered a System Usability Scale (SUS) \cite{brooke1996sus} to estimate the usability of the guidance system. Finally, we conducted a semi-structured interview asking what participants liked and disliked about the system, what needed to be improved, and how they might envision an ideal guidance system. Since P4 is a co-author on this paper, we did not administer the SUS or semi-structured interview to him.

# Experimental Results and Analysis {#results}

We conducted an experiment to assess the feasibility of the spatial guidance feedback and to determine whether the feedback speeds up the target search. The experiment (\ref{fig:exp}) compared how much time a blind participant needed to find hotspots (points of interest) on a 3D object with spatial guidance vs. without it. The object was a large toy fire truck (27 cm x 12 cm x 14 cm), with K=10 distinct hotspot locations defined in advance. The hotspots were tactilely salient features such as lights, hooks or other features that protruded from the truck. Some hotspots were in recessed locations on the truck, such as the shift stick and steering wheel (\ref{fig:exp}) in the front cabin, or on the side of the truck facing away from the participant; these hotspots were difficult to find and access, whether by fingertip or by stylus. In fact, the difficulty of accessing such locations highlights an advantage of the CamIO stylus, whose tip need not be visible to the camera. This allows the user to explore recessed areas or areas that are otherwise invisible to the camera.

\begin{figure}
\includegraphics[width=5.5cm, keepaspectratio,]{res/img/fig2a.jpg}
\includegraphics[width=6.5cm, keepaspectratio,]{res/img/fig2b.jpg}
\caption{(left) Plot of time (x axis) to find target hotspot location for G (‘o’ symbol) vs. NG  (‘x’ symbol) conditions, with time-out declared after 120 sec. of unsuccessful search. Results are shown for each participant (indicated on the y axis and also using color) and at all 10 target locations. (right) Histogram plot showing the distribution of NG time minus G time. The plot aggregates time differences for all participants over all locations; positive times mean the search was faster with G than with NG.}
\label{fig:res}
\end{figure}

We analyzed the differences between times under G vs. NG using a Wilcoxon signed rank test. The resulting p-value is p = 3.8 x 10-8. This is a highly statistically significant p-value supporting the rejection of the null hypothesis that the distribution of times under G is the same as the distribution of times under NG.

The following System Usability Scale (SUS) scores were calculated: P1=65, P2=75, P3=77.5, for an average of 72.5. An SUS score of 68 is considered “average” while scores in the mid- to high 70s are considered “good” (\url{https://measuringu.com/interpret-sus-score/}). Overall, the SUS scores suggest that the guidance system is usable but needs improvement.

Next we summarize the qualitative feedback we obtained in the semi-structured interviews. All participants who provided qualitative feedback (P1, P2 and P3) acknowledged the benefit of the audio guidance that was provided, which they found intuitive and clearly superior to finding hotspots without guidance. However, P1 and P2 felt that the limited spatial accuracy of the CamIO system was the biggest problem that needed to be addressed; for instance, in some cases the system announced the target had been reached even though the stylus tip was hovering about 1 cm above the truck surface. (This spatial inaccuracy results from noise in the board pose and stylus tip location estimates.) P3 expressed frustration with time lag, i.e., the voice feedback tended to lag behind the current location of the stylus, even when she was moving the stylus slowly; she also felt that the speed of the repetitive speech feedback made her feel rushed.

When asked what improvements the participants wanted, P1 suggested equipping the stylus with a button that turns on the guidance feedback, and the addition of multiple vibrators on the stylus that could provide haptic feedback indicating which way the stylus should be moved. P2 suggested adding audio information similar to the earlier guidance system that the authors experimented with (Sec. 3), which would use a combination of tempo, pitch and volume changes to signal when the stylus gets closer to the target location. P3 expressed the desire that CamIO work with any ordinary pen or pencil, rather than requiring the special stylus we used.

# Conclusion {#conclusion}

We have devised a 3D spatial guidance system that helps a blind user find a specific location of interest on an object using CamIO, an audio-based AR system. The guidance system uses simple verbal commands (“left”/ “right”/“up”/etc.) to tell the user which direction to move the stylus tip. Experiments with four blind participants show that the guidance system significantly speeds up the search for target hotspots compared with an exhaustive search. A secondary contribution of this study is to demonstrate the ability of the CamIO stylus to access recessed or otherwise obscured locations on an object that are not themselves visible to the camera.

While we applied this guidance approach specifically to CamIO, it could be useful in any AR system, whether the system tracks the user’s hands, handheld stylus or other pointing tool. Such audio-based guidance is not only helpful for blind users but is also useful in applications where visibility is limited (either by environmental conditions or because of a visual impairment), or for sighted users who prefer a cross-sensory display (e.g., audio and visual combined).

Future work will focus on refining and improving the audio interface. One possible way to speed up the guidance process is to add more specific directional feedback, e.g., “far left” would tell the user to move the stylus farther to the left than the “left” instruction. The verbal directions could be augmented with 3D spatialized sound as in \cite{mayAuditoryDisplaysFacilitate2019}, which could make the guidance more intuitive. Finally, we will continue to test and improve the CamIO system, with an emphasis on refining the spatial accuracy of the tip location estimates using improved computer vision algorithms.

### 2020-05-28 Manuscript.tex
Init

--- 2020-05-28 Manuscript.Rmd
@@
----
-title: "An Audio-Based 3D Spatial Guidance AR System for Blind Users"
-titlerunning: Audio-Based 3D Spatial Guidance
-authorrunning: J. Couhlan et al.
-output:
-  pdf_document:
-    template: res/templates/markdown.tex
-    toc: no
-    keep_tex: yes
-institutes:
-- id: 1
-  name: The Smith-Kettlewell Eye Research Institute, San Francisco, CA
-  emails: coughlan@ski.org, brandon.biggs@ski.org, hshen@ski.org
-- id: 2
-  name: LITIS, University of Rouen, Normandy, France
-  emails: marc-aurele.riviere@univ-rouen.fr
-keywords:
-- Assisitive Devices
-- Augmented Reality
-- Auditory Substitution
-- Visual Impairment
-abstract: Augmented reality (AR) has great potential for blind users because it enables
-  a range of applications that provide audio information about specific locations
-  or directions in the user’s environment. For instance, the CamIO (“Camera Input-Output”)
-  AR app makes physical objects (such as documents, maps, devices and 3D models) accessible
-  to blind and visually impaired persons by providing real-time audio feedback in
-  response to the location on an object that the user is touching (using an inexpensive
-  stylus). An important feature needed by blind users of AR apps such as CamIO is
-  a 3D spatial guidance feature that provides real-time audio feedback to help the
-  user find a desired location on an object. We have devised a simple audio interface
-  to provide verbal guidance towards a target of interest in 3D. The experiment we
-  report with blind participants using this guidance interface demonstrates the feasibility
-  of the approach and its benefit for helping users find locations of interest.
-authors:
-- name: James Coughlan
-  orcid: 0000-1111-2222-3333
-  inst: 1
-- name: Brandon Biggs
-  orcid: 0000-1111-2222-3333
-  inst: 1
-- name: Marc-Aurèle Rivière
-  orcid: 0000-0002-5108-3382
-  inst: 2
-- name: Huiying Shen
-  orcid: 0000-1111-2222-3333
-  inst: 1
-ressources: res/
-templates: res/templates/
-bibliography: res/refs.bib
----
+% This is template.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.20 of 2017/10/04
 
-```{r setup, include=FALSE}
+\documentclass[runningheads]{res/templates/llncs}
 
-#TODO: DESCRIPTION with package metadata ?
+\usepackage{graphicx}
+\usepackage[utf8]{inputenc}
+\usepackage[hyphens]{url}
+\usepackage{hyperref}
+\usepackage{etoolbox}
 
-knitr::opts_chunk$set(
-  fig.show="hold",
-  fig.align="default",
-  warning = FALSE,
-  echo = FALSE,
-  cache = FALSE,
-  error = FALSE
-)
+% Include packages from YAML header
 
-colorize <- function(x, color) {
-  if (knitr::is_latex_output()) {
-    sprintf("\\textcolor{%s}{%s}", color, x)
-  } else if (knitr::is_html_output()) {
-    sprintf("<span style='color: %s;'>%s</span>", color, x)
-  } else x
+\begin{document}
+
+\title{An Audio-Based 3D Spatial Guidance AR System for Blind Users}
+
+\titlerunning{Audio-Based 3D Spatial Guidance}
+%
+\newbool{flag}
+\boolfalse{flag}
+%
+\author{
+\ifbool{flag}{\and}{\booltrue{flag}}James Coughlan\inst{1}\orcidID{0000-1111-2222-3333}
+\ifbool{flag}{\and}{\booltrue{flag}}Brandon Biggs\inst{1}\orcidID{0000-1111-2222-3333}
+\ifbool{flag}{\and}{\booltrue{flag}}Marc-Aurèle Rivière\inst{2}\orcidID{0000-0002-5108-3382}
+\ifbool{flag}{\and}{\booltrue{flag}}Huiying Shen\inst{1}\orcidID{0000-1111-2222-3333}
 }
-```
+%
+\authorrunning{J. Couhlan et al.}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+\boolfalse{flag}
+%
+\institute{
+\ifbool{flag}{\and}{\booltrue{flag}}
+ The Smith-Kettlewell Eye Research Institute, San Francisco, CA \\
+ \email{\href{mailto:coughlan@ski.org}{\nolinkurl{coughlan@ski.org}},
+\href{mailto:brandon.biggs@ski.org}{\nolinkurl{brandon.biggs@ski.org}},
+\href{mailto:hshen@ski.org}{\nolinkurl{hshen@ski.org}}} 
+\ifbool{flag}{\and}{\booltrue{flag}}
+ LITIS, University of Rouen, Normandy, France \\
+ \email{\href{mailto:marc-aurele.riviere@univ-rouen.fr}{\nolinkurl{marc-aurele.riviere@univ-rouen.fr}}} 
+}
 
-# State of the Art and Related Technology {#sota}
+%
+\maketitle              % typeset the header of the contribution
+%
+\begin{abstract}
+Augmented reality (AR) has great potential for blind users because it
+enables a range of applications that provide audio information about
+specific locations or directions in the user's environment. For
+instance, the CamIO (``Camera Input-Output'') AR app makes physical
+objects (such as documents, maps, devices and 3D models) accessible to
+blind and visually impaired persons by providing real-time audio
+feedback in response to the location on an object that the user is
+touching (using an inexpensive stylus). An important feature needed by
+blind users of AR apps such as CamIO is a 3D spatial guidance feature
+that provides real-time audio feedback to help the user find a desired
+location on an object. We have devised a simple audio interface to
+provide verbal guidance towards a target of interest in 3D. The
+experiment we report with blind participants using this guidance
+interface demonstrates the feasibility of the approach and its benefit
+for helping users find locations of interest.
+\\
+\boolfalse{flag}
+\keywords{
+\ifbool{flag}{\and}{\booltrue{flag}}
+Assisitive Devices
+\ifbool{flag}{\and}{\booltrue{flag}}
+Augmented Reality
+\ifbool{flag}{\and}{\booltrue{flag}}
+Auditory Substitution
+\ifbool{flag}{\and}{\booltrue{flag}}
+Visual Impairment
+}
+\end{abstract}
+%
+%
+\hypertarget{sota}{%
+\section{State of the Art and Related Technology}\label{sota}}
 
-Many people who are blind or visually impaired have difficulties accessing a range of everyday objects, including printed documents, maps, infographics, appliances and 3D models used in STEM education, needed for daily activities in schools, the home and the workplace.
-This limits their participation in many cultural, professional and educational activities.
-While braille labeling is often a useful means of providing access to such objects, there is only limited space for braille, and this method is only accessible to those who can read braille.
+Many people who are blind or visually impaired have difficulties
+accessing a range of everyday objects, including printed documents,
+maps, infographics, appliances and 3D models used in STEM education,
+needed for daily activities in schools, the home and the workplace. This
+limits their participation in many cultural, professional and
+educational activities. While braille labeling is often a useful means
+of providing access to such objects, there is only limited space for
+braille, and this method is only accessible to those who can read
+braille.
 
 Audio labels are a powerful supplement or alternative to braille labels.
-Standard methods of implementing audio labels require the use of special hardware and materials, which is costly and limits their adoption.
-For instance, tactile graphics may be overlaid on a touch-sensitive tablet \cite{TalkingTactileTablet}.
-Documents or other surfaces can be covered with a special form of paper readable by a “smart pen” \cite{mieleTalkingTactileApps2010}.
-The PenFriend \cite{kendrickPenFriendTouchMemo2011} stylus uses small barcodes affixed to an object to define audio labels.
+Standard methods of implementing audio labels require the use of special
+hardware and materials, which is costly and limits their adoption. For
+instance, tactile graphics may be overlaid on a touch-sensitive tablet
+\cite{TalkingTactileTablet}. Documents or other surfaces can be covered
+with a special form of paper readable by a ``smart pen''
+\cite{mieleTalkingTactileApps2010}. The PenFriend
+\cite{kendrickPenFriendTouchMemo2011} stylus uses small barcodes affixed
+to an object to define audio labels.
 
-There is growing interest in using computer vision-based approaches for audio labeling.
-In these approaches a camera tracks the user’s fingers or handheld pointer as they explore an object, enabling audio labeling for existing objects with minimal or no customization.
-Past computer vision approaches to audio labeling include \cite{shiMarkitTalkitLowBarrier2017}, which focuses on 3D printed objects, \cite{thevinAugmentedRealityPeople2018}, which uses a depth camera to access flat documents, and \cite{fuscoTactileGraphicsHelper2015}, which provides access to tactile graphics to students in educational settings.
+There is growing interest in using computer vision-based approaches for
+audio labeling. In these approaches a camera tracks the user's fingers
+or handheld pointer as they explore an object, enabling audio labeling
+for existing objects with minimal or no customization. Past computer
+vision approaches to audio labeling include
+\cite{shiMarkitTalkitLowBarrier2017}, which focuses on 3D printed
+objects, \cite{thevinAugmentedRealityPeople2018}, which uses a depth
+camera to access flat documents, and
+\cite{fuscoTactileGraphicsHelper2015}, which provides access to tactile
+graphics to students in educational settings.
 
-We build on our past work on CamIO \cite{shenCamIO3DComputer2013,coughlanEvaluatingAuthorUser2017,coughlanTowardsAccessibleAudioLabeling2020}, which is short for “Camera Input-Output.” We have implemented CamIO as a stand-alone iPhone app that tracks an inexpensive wooden stylus held by the user to point at locations of interest on an object (called hotspots).
-When the user points at a hotspot with the stylus tip, relevant audio information, such as text-to-speech or other sounds, is announced.
+We build on our past work on CamIO
+\cite{shenCamIO3DComputer2013,coughlanEvaluatingAuthorUser2017,coughlanTowardsAccessibleAudioLabeling2020},
+which is short for ``Camera Input-Output.'' We have implemented CamIO as
+a stand-alone iPhone app that tracks an inexpensive wooden stylus held
+by the user to point at locations of interest on an object (called
+hotspots). When the user points at a hotspot with the stylus tip,
+relevant audio information, such as text-to-speech or other sounds, is
+announced.
 
-Experiments with blind participants have demonstrated the feasibility of using CamIO to access existing audio labels.
-Moreover, our approach allows a visually impaired user to create audio labels independently \cite{coughlanTowardsAccessibleAudioLabeling2020}.
-However, an ability that CamIO still lacks is the ability to provide real-time guidance to a hotspot.
-Without such guidance, the user must search the entire surface of an object with the stylus tip until they hear the hotspot they are seeking.
-This need is particularly acute in cases where the surface is lacking in tactile cues, such as the touch screen buttons on a microwave oven \cite{coughlanTowardsAccessibleAudioLabeling2020}.
+Experiments with blind participants have demonstrated the feasibility of
+using CamIO to access existing audio labels. Moreover, our approach
+allows a visually impaired user to create audio labels independently
+\cite{coughlanTowardsAccessibleAudioLabeling2020}. However, an ability
+that CamIO still lacks is the ability to provide real-time guidance to a
+hotspot. Without such guidance, the user must search the entire surface
+of an object with the stylus tip until they hear the hotspot they are
+seeking. This need is particularly acute in cases where the surface is
+lacking in tactile cues, such as the touch screen buttons on a microwave
+oven \cite{coughlanTowardsAccessibleAudioLabeling2020}.
 
-Guidance not only saves the user time in exploring large or intricate objects, or objects lacking in tactile features, but can enable entirely new ways of interacting with objects.
-For instance, a geography learning app could ask a student to point to a specific city on a world globe, and if the student has trouble finding it, the app could guide them to it in an intuitive way.
-Other examples could include guiding a user to a specific button on an appliance (such as a microwave) to perform a desired function, guiding a student to a specific element on the periodic table, or even guiding a repairperson to find a part of a machine with limited visibility.
-Our guidance system is inspired by recent work on auditory and haptic displays for 3D spatial guidance \cite{mayAuditoryDisplaysFacilitate2019,guezou-philippePrototypingEvaluatingSensory2018}, which provide real-time audio/haptic feedback to guide a user towards a target in a virtual environment.
-We initially developed an interface similar to \cite{mayAuditoryDisplaysFacilitate2019} for CamIO, but after preliminary testing we devised a simpler and more effective interface for our application, which we evaluated experimentally with four blind participants.
+Guidance not only saves the user time in exploring large or intricate
+objects, or objects lacking in tactile features, but can enable entirely
+new ways of interacting with objects. For instance, a geography learning
+app could ask a student to point to a specific city on a world globe,
+and if the student has trouble finding it, the app could guide them to
+it in an intuitive way. Other examples could include guiding a user to a
+specific button on an appliance (such as a microwave) to perform a
+desired function, guiding a student to a specific element on the
+periodic table, or even guiding a repairperson to find a part of a
+machine with limited visibility. Our guidance system is inspired by
+recent work on auditory and haptic displays for 3D spatial guidance
+\cite{mayAuditoryDisplaysFacilitate2019,guezou-philippePrototypingEvaluatingSensory2018},
+which provide real-time audio/haptic feedback to guide a user towards a
+target in a virtual environment. We initially developed an interface
+similar to \cite{mayAuditoryDisplaysFacilitate2019} for CamIO, but after
+preliminary testing we devised a simpler and more effective interface
+for our application, which we evaluated experimentally with four blind
+participants.
 
-# Overview of the CamIO System {#overview}
+\hypertarget{overview}{%
+\section{Overview of the CamIO System}\label{overview}}
 
-CamIO is an augmented reality (AR) iPhone app that uses computer vision to provide audio labels for rigid 3D objects.
-The rigid object of interest is mounted on a board, which is a printed barcode pattern (\ref{fig:exp}) that allows CamIO to estimate the object’s pose (3D translation and 3D orientation) relative to the camera.
-The iPhone is typically mounted on a tripod or a gooseneck cellphone holder \ref{fig:exp} to obtain a clear view of the board and object, but in some cases \cite{coughlanTowardsAccessibleAudioLabeling2020} users may hold the iPhone by hand.
-The user points the tip of a passive wooden stylus \ref{fig:exp}, covered with another barcode pattern, to different locations on an object.
-CamIO estimates the stylus tip’s 3D location relative to the camera and uses the pose estimate of the board to determine the stylus tip’s 3D location relative to the object.
-Whenever the tip is close enough to a pre-defined hotspot, it triggers audio information about the hotspot.
-(In other words, hotspot locations are defined relative to the board, which is mounted rigidly to the object.)
+CamIO is an augmented reality (AR) iPhone app that uses computer vision
+to provide audio labels for rigid 3D objects. The rigid object of
+interest is mounted on a board, which is a printed barcode pattern
+(\ref{fig:exp}) that allows CamIO to estimate the object's pose (3D
+translation and 3D orientation) relative to the camera. The iPhone is
+typically mounted on a tripod or a gooseneck cellphone holder
+\ref{fig:exp} to obtain a clear view of the board and object, but in
+some cases \cite{coughlanTowardsAccessibleAudioLabeling2020} users may
+hold the iPhone by hand. The user points the tip of a passive wooden
+stylus \ref{fig:exp}, covered with another barcode pattern, to different
+locations on an object. CamIO estimates the stylus tip's 3D location
+relative to the camera and uses the pose estimate of the board to
+determine the stylus tip's 3D location relative to the object. Whenever
+the tip is close enough to a pre-defined hotspot, it triggers audio
+information about the hotspot. (In other words, hotspot locations are
+defined relative to the board, which is mounted rigidly to the object.)
 
-A second stylus (not shown, identical in form but distinct from the one shown in \ref{fig:exp}) is used to create new hotspots, and allows the user to make an audio recording to associate with each new hotspot.
-Our recent work \cite{coughlanTowardsAccessibleAudioLabeling2020} demonstrates the accessibility of this authoring process, which allows blind users to create their own annotations (hotspots and associated audio labels) independently.
-However, in the study presented in this paper, the experimenter was the only one who created hotspots, since the emphasis here is on providing guidance to pre-existing hotspots.
+A second stylus (not shown, identical in form but distinct from the one
+shown in \ref{fig:exp}) is used to create new hotspots, and allows the
+user to make an audio recording to associate with each new hotspot. Our
+recent work \cite{coughlanTowardsAccessibleAudioLabeling2020}
+demonstrates the accessibility of this authoring process, which allows
+blind users to create their own annotations (hotspots and associated
+audio labels) independently. However, in the study presented in this
+paper, the experimenter was the only one who created hotspots, since the
+emphasis here is on providing guidance to pre-existing hotspots.
 
-# 3D Spatial Guidance Interface {#material}
+\hypertarget{material}{%
+\section{3D Spatial Guidance Interface}\label{material}}
 
-We build on the auditory display described in \cite{mayAuditoryDisplaysFacilitate2019}, which is intended for audio-based 3D targeting of locations in a large virtual workspace (such as a kitchen).
-Accordingly, the first incarnation of our guidance approach used a repeating monaural sound, whose tempo indicated how far the stylus tip is from the target, and with a pitch (low, medium or high) that indicated whether the tip needed to move up or down (or else was aligned properly in the vertical direction).
+We build on the auditory display described in
+\cite{mayAuditoryDisplaysFacilitate2019}, which is intended for
+audio-based 3D targeting of locations in a large virtual workspace (such
+as a kitchen). Accordingly, the first incarnation of our guidance
+approach used a repeating monaural sound, whose tempo indicated how far
+the stylus tip is from the target, and with a pitch (low, medium or
+high) that indicated whether the tip needed to move up or down (or else
+was aligned properly in the vertical direction).
 
-Informal tests with two blind participants (one female of age 42, who is also participant P1 in the experiment described in the next section, and one male of age 73) showed that this guidance worked satisfactorily in some cases.
-However, we found that the feedback was too indirect: the participants didn’t always know how to move the stylus to increase the tempo.
-As a result, the resulting guidance process was unacceptably slow.
-Thus we devised a second, more direct guidance approach: verbal directions telling the user to move the stylus along the cardinal directions (up, down, left, right, forward or back).
-While the optimal direction to move the stylus tip towards the hotspot might entail a combination of these directions (e.g., simultaneously move left and up), for simplicity we issue directions along only one cardinal direction at a time – the direction that will most improve the alignment to the target.
+Informal tests with two blind participants (one female of age 42, who is
+also participant P1 in the experiment described in the next section, and
+one male of age 73) showed that this guidance worked satisfactorily in
+some cases. However, we found that the feedback was too indirect: the
+participants didn't always know how to move the stylus to increase the
+tempo. As a result, the resulting guidance process was unacceptably
+slow. Thus we devised a second, more direct guidance approach: verbal
+directions telling the user to move the stylus along the cardinal
+directions (up, down, left, right, forward or back). While the optimal
+direction to move the stylus tip towards the hotspot might entail a
+combination of these directions (e.g., simultaneously move left and up),
+for simplicity we issue directions along only one cardinal direction at
+a time -- the direction that will most improve the alignment to the
+target.
 
-We define the guidance interface as follows.
-The board (barcode pattern beneath the truck, see \ref{fig:exp}) defines an xyz coordinate frame, with $+x$ = right, $-x$ = left, $+y$ = forward (away from the user), $-y$ = back (towards the user), $+z$ = up and $-z$ = down.
-Let the desired target hotspot location be denoted $(x*, y*, z*)$, and the current location of the stylus tip be $(x,y,z)$.
-Define the error vector $e = (x-x*, y-y*, z-z*)$, and let index $a = argmax_i |e_i|$, i.e., a is the index (1, 2 or 3, corresponding to x, y or z, respectively) of the entry in e with the highest absolute value.
-In other words, a indicates which coordinate (x, y or z) is most discrepant from the target coordinate location $(x*, y*, z*)$.
-Then the appropriate direction is issued, e.g., “left” if $a = 1$ and $x > x*$, “right” if $a = 1$ and $x < x*$, “up” if $a = 3$ and $z < z*$, etc.
-Directions are announced roughly twice per second whenever the stylus is visible.
-The target hotspot is announced when the stylus gets to within approximately 1 cm from the hotspot.
-No audio feedback is issued when the stylus is not visible.
+We define the guidance interface as follows. The board (barcode pattern
+beneath the truck, see \ref{fig:exp}) defines an xyz coordinate frame,
+with \(+x\) = right, \(-x\) = left, \(+y\) = forward (away from the
+user), \(-y\) = back (towards the user), \(+z\) = up and \(-z\) = down.
+Let the desired target hotspot location be denoted \((x*, y*, z*)\), and
+the current location of the stylus tip be \((x,y,z)\). Define the error
+vector \(e = (x-x*, y-y*, z-z*)\), and let index \(a = argmax_i |e_i|\),
+i.e., a is the index (1, 2 or 3, corresponding to x, y or z,
+respectively) of the entry in e with the highest absolute value. In
+other words, a indicates which coordinate (x, y or z) is most discrepant
+from the target coordinate location \((x*, y*, z*)\). Then the
+appropriate direction is issued, e.g., ``left'' if \(a = 1\) and
+\(x > x*\), ``right'' if \(a = 1\) and \(x < x*\), ``up'' if \(a = 3\)
+and \(z < z*\), etc. Directions are announced roughly twice per second
+whenever the stylus is visible. The target hotspot is announced when the
+stylus gets to within approximately 1 cm from the hotspot. No audio
+feedback is issued when the stylus is not visible.
 
-# Experiment with Blind Participants {#exp}
+\hypertarget{exp}{%
+\section{Experiment with Blind Participants}\label{exp}}
 
-We conducted an experiment to assess the feasibility of the spatial guidance feedback and to determine whether the feedback speeds up the target search.
-The experiment \ref{fig:exp} compared how much time a blind participant needed to find hotspots (points of interest) on a 3D object with spatial guidance vs.
-without it.
-The object was a large toy fire truck (27 cm x 12 cm x 14 cm), with K=10 distinct hotspot locations defined in advance.
-The hotspots were tactilely salient features such as lights, hooks or other features that protruded from the truck.
-Some hotspots were in recessed locations on the truck, such as the shift stick and steering wheel (\ref{fig:exp}) in the front cabin, or on the side of the truck facing away from the participant; these hotspots were difficult to find and access, whether by fingertip or by stylus.
-In fact, the difficulty of accessing such locations highlights an advantage of the CamIO stylus, whose tip need not be visible to the camera.
-This allows the user to explore recessed areas or areas that are otherwise invisible to the camera.
+We conducted an experiment to assess the feasibility of the spatial
+guidance feedback and to determine whether the feedback speeds up the
+target search. The experiment \ref{fig:exp} compared how much time a
+blind participant needed to find hotspots (points of interest) on a 3D
+object with spatial guidance vs.~without it. The object was a large toy
+fire truck (27 cm x 12 cm x 14 cm), with K=10 distinct hotspot locations
+defined in advance. The hotspots were tactilely salient features such as
+lights, hooks or other features that protruded from the truck. Some
+hotspots were in recessed locations on the truck, such as the shift
+stick and steering wheel (\ref{fig:exp}) in the front cabin, or on the
+side of the truck facing away from the participant; these hotspots were
+difficult to find and access, whether by fingertip or by stylus. In
+fact, the difficulty of accessing such locations highlights an advantage
+of the CamIO stylus, whose tip need not be visible to the camera. This
+allows the user to explore recessed areas or areas that are otherwise
+invisible to the camera.
 
-```{r exp, out.width=c("55%","45%"), fig.cap="(left) Blind participant uses CamIO to explore a toy fire truck.
+\begin{figure}
+\includegraphics[width=0.55\linewidth]{res/img/fig1a} \includegraphics[width=0.45\linewidth]{res/img/fig1b} \caption{(left) Blind participant uses CamIO to explore a toy fire truck.
 The participant holds the CamIO stylus (a barcode pattern printed on a square wooden dowel with a pointy tip) to point to possible hotspots on the truck.
@@
 (right) Stylus points to steering wheel in the truck cabin.
-Though the steering wheel is in a recessed location that is not visible to CamIO, the stylus tip location can still be estimated."}
+Though the steering wheel is in a recessed location that is not visible to CamIO, the stylus tip location can still be estimated.}\label{fig:exp}
+\end{figure}
 
-knitr::include_graphics("res/img/fig1a.jpg")
-knitr::include_graphics("res/img/fig1b.jpg")
-```
+Each hotspot location was searched for under two feedback conditions: G
+(guidance provided) and NG (no guidance). The participant was given up
+to 120 sec.~to find each hotspot, after which a time-out was declared.
+In order to minimize the influence of any possible learning effects,
+which could make it easier to find a hotspot the second time it was
+searched for, we randomized the sequence of hotspots and alternated
+between NG and G trials. This resulted in a sequence of 2K = 20 trials,
+starting with an NG trial, the same sequence for each participant.
 
-Each hotspot location was searched for under two feedback conditions: G (guidance provided) and NG (no guidance).
-The participant was given up to 120 sec.
-to find each hotspot, after which a time-out was declared.
-In order to minimize the influence of any possible learning effects, which could make it easier to find a hotspot the second time it was searched for, we randomized the sequence of hotspots and alternated between NG and G trials.
-This resulted in a sequence of 2K = 20 trials, starting with an NG trial, the same sequence for each participant.
+We had a total of N=4 blind participants in our experiment (2 male/2
+female, ages from 26-42). Participants P1, P3 and P4 were already
+familiar with the CamIO system (P4 is a co-author of this paper) while
+P2 was unfamiliar with it. None of the participants was told the study
+design before their trials, and we didn't tell them whether or not any
+locations would be repeated in the 20 trials. After obtaining IRB
+consent, we reviewed how CamIO works if necessary and demonstrated the
+audio guidance feature. The experimenter launched CamIO on an iPhone 8
+and positioned it on a gooseneck cellphone holder so that the camera
+captured the entire truck and board. Each participant spent a short time
+practicing with the guidance feature, with help provided by the
+experimenter as needed. (The experimenter chose practice hotspots that
+were different from the pre-defined K=10 hotspots.) In the practice
+phase, the experimenter explained that barriers sometimes block the user
+from moving in the suggested guidance direction. For instance, if in
+\ref{fig:exp} the system directs the participant to move ``forward''
+towards the left side mirror, the stylus will soon hit the cabin
+roof/wall as it moves towards the mirror; the participant must find ways
+around such obstacles to find the target hotspot.
 
-We had a total of N=4 blind participants in our experiment (2 male/2 female, ages from 26-42).
-Participants P1, P3 and P4 were already familiar with the CamIO system (P4 is a co-author of this paper) while P2 was unfamiliar with it.
-None of the participants was told the study design before their trials, and we didn’t tell them whether or not any locations would be repeated in the 20 trials.
-After obtaining IRB consent, we reviewed how CamIO works if necessary and demonstrated the audio guidance feature.
-The experimenter launched CamIO on an iPhone 8 and positioned it on a gooseneck cellphone holder so that the camera captured the entire truck and board.
-Each participant spent a short time practicing with the guidance feature, with help provided by the experimenter as needed.
-(The experimenter chose practice hotspots that were different from the pre-defined K=10 hotspots.) In the practice phase, the experimenter explained that barriers sometimes block the user from moving in the suggested guidance direction.
-For instance, if in \ref{fig:exp} the system directs the participant to move “forward” towards the left side mirror, the stylus will soon hit the cabin roof/wall as it moves towards the mirror; the participant must find ways around such obstacles to find the target hotspot.
+Next each participant performed the formal experimental trials, which
+consisted of the 2K=20 trial sequence described above. This sequence
+included 10 trials of both conditions (G and NG), for a total of 20
+trials per participant. After completing the 20 search trials, we
+administered a System Usability Scale (SUS) \cite{brooke1996sus} to
+estimate the usability of the guidance system. Finally, we conducted a
+semi-structured interview asking what participants liked and disliked
+about the system, what needed to be improved, and how they might
+envision an ideal guidance system. Since P4 is a co-author on this
+paper, we did not administer the SUS or semi-structured interview to
+him.
 
-Next each participant performed the formal experimental trials, which consisted of the 2K=20 trial sequence described above.
-This sequence included 10 trials of both conditions (G and NG), for a total of 20 trials per participant.
-After completing the 20 search trials, we administered a System Usability Scale (SUS) \cite{brooke1996sus} to estimate the usability of the guidance system.
-Finally, we conducted a semi-structured interview asking what participants liked and disliked about the system, what needed to be improved, and how they might envision an ideal guidance system.
-Since P4 is a co-author on this paper, we did not administer the SUS or semi-structured interview to him.
+\hypertarget{results}{%
+\section{Experimental Results and Analysis}\label{results}}
 
-# Experimental Results and Analysis {#results}
-
-We conducted an experiment to assess the feasibility of the spatial guidance feedback and to determine whether the feedback speeds up the target search.
-The experiment (\ref{fig:exp}) compared how much time a blind participant needed to find hotspots (points of interest) on a 3D object with spatial guidance vs.
-without it.
-The object was a large toy fire truck (27 cm x 12 cm x 14 cm), with K=10 distinct hotspot locations defined in advance.
-The hotspots were tactilely salient features such as lights, hooks or other features that protruded from the truck.
-Some hotspots were in recessed locations on the truck, such as the shift stick and steering wheel (\ref{fig:exp}) in the front cabin, or on the side of the truck facing away from the participant; these hotspots were difficult to find and access, whether by fingertip or by stylus.
-In fact, the difficulty of accessing such locations highlights an advantage of the CamIO stylus, whose tip need not be visible to the camera.
-This allows the user to explore recessed areas or areas that are otherwise invisible to the camera.
+We conducted an experiment to assess the feasibility of the spatial
+guidance feedback and to determine whether the feedback speeds up the
+target search. The experiment (\ref{fig:exp}) compared how much time a
+blind participant needed to find hotspots (points of interest) on a 3D
+object with spatial guidance vs.~without it. The object was a large toy
+fire truck (27 cm x 12 cm x 14 cm), with K=10 distinct hotspot locations
+defined in advance. The hotspots were tactilely salient features such as
+lights, hooks or other features that protruded from the truck. Some
+hotspots were in recessed locations on the truck, such as the shift
+stick and steering wheel (\ref{fig:exp}) in the front cabin, or on the
+side of the truck facing away from the participant; these hotspots were
+difficult to find and access, whether by fingertip or by stylus. In
+fact, the difficulty of accessing such locations highlights an advantage
+of the CamIO stylus, whose tip need not be visible to the camera. This
+allows the user to explore recessed areas or areas that are otherwise
+invisible to the camera.
 
@@
 
-We analyzed the differences between times under G vs.
-NG using a Wilcoxon signed rank test.
-The resulting p-value is p = 3.8 x 10-8.
-This is a highly statistically significant p-value supporting the rejection of the null hypothesis that the distribution of times under G is the same as the distribution of times under NG.
+We analyzed the differences between times under G vs.~NG using a
+Wilcoxon signed rank test. The resulting p-value is p = 3.8 x 10-8. This
+is a highly statistically significant p-value supporting the rejection
+of the null hypothesis that the distribution of times under G is the
+same as the distribution of times under NG.
 
-The following System Usability Scale (SUS) scores were calculated: P1=65, P2=75, P3=77.5, for an average of 72.5.
-An SUS score of 68 is considered “average” while scores in the mid- to high 70s are considered “good” (\url{https://measuringu.com/interpret-sus-score/}).
-Overall, the SUS scores suggest that the guidance system is usable but needs improvement.
+The following System Usability Scale (SUS) scores were calculated:
+P1=65, P2=75, P3=77.5, for an average of 72.5. An SUS score of 68 is
+considered ``average'' while scores in the mid- to high 70s are
+considered ``good'' (\url{https://measuringu.com/interpret-sus-score/}).
+Overall, the SUS scores suggest that the guidance system is usable but
+needs improvement.
 
-Next we summarize the qualitative feedback we obtained in the semi-structured interviews.
-All participants who provided qualitative feedback (P1, P2 and P3) acknowledged the benefit of the audio guidance that was provided, which they found intuitive and clearly superior to finding hotspots without guidance.
-However, P1 and P2 felt that the limited spatial accuracy of the CamIO system was the biggest problem that needed to be addressed; for instance, in some cases the system announced the target had been reached even though the stylus tip was hovering about 1 cm above the truck surface.
-(This spatial inaccuracy results from noise in the board pose and stylus tip location estimates.) P3 expressed frustration with time lag, i.e., the voice feedback tended to lag behind the current location of the stylus, even when she was moving the stylus slowly; she also felt that the speed of the repetitive speech feedback made her feel rushed.
+Next we summarize the qualitative feedback we obtained in the
+semi-structured interviews. All participants who provided qualitative
+feedback (P1, P2 and P3) acknowledged the benefit of the audio guidance
+that was provided, which they found intuitive and clearly superior to
+finding hotspots without guidance. However, P1 and P2 felt that the
+limited spatial accuracy of the CamIO system was the biggest problem
+that needed to be addressed; for instance, in some cases the system
+announced the target had been reached even though the stylus tip was
+hovering about 1 cm above the truck surface. (This spatial inaccuracy
+results from noise in the board pose and stylus tip location estimates.)
+P3 expressed frustration with time lag, i.e., the voice feedback tended
+to lag behind the current location of the stylus, even when she was
+moving the stylus slowly; she also felt that the speed of the repetitive
+speech feedback made her feel rushed.
 
-When asked what improvements the participants wanted, P1 suggested equipping the stylus with a button that turns on the guidance feedback, and the addition of multiple vibrators on the stylus that could provide haptic feedback indicating which way the stylus should be moved.
-P2 suggested adding audio information similar to the earlier guidance system that the authors experimented with (Sec.
-3), which would use a combination of tempo, pitch and volume changes to signal when the stylus gets closer to the target location.
-P3 expressed the desire that CamIO work with any ordinary pen or pencil, rather than requiring the special stylus we used.
+When asked what improvements the participants wanted, P1 suggested
+equipping the stylus with a button that turns on the guidance feedback,
+and the addition of multiple vibrators on the stylus that could provide
+haptic feedback indicating which way the stylus should be moved. P2
+suggested adding audio information similar to the earlier guidance
+system that the authors experimented with (Sec. 3), which would use a
+combination of tempo, pitch and volume changes to signal when the stylus
+gets closer to the target location. P3 expressed the desire that CamIO
+work with any ordinary pen or pencil, rather than requiring the special
+stylus we used.
 
-# Conclusion {#conclusion}
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 
-We have devised a 3D spatial guidance system that helps a blind user find a specific location of interest on an object using CamIO, an audio-based AR system.
-The guidance system uses simple verbal commands (“left”/ “right”/“up”/etc.) to tell the user which direction to move the stylus tip.
-Experiments with four blind participants show that the guidance system significantly speeds up the search for target hotspots compared with an exhaustive search.
-A secondary contribution of this study is to demonstrate the ability of the CamIO stylus to access recessed or otherwise obscured locations on an object that are not themselves visible to the camera.
+We have devised a 3D spatial guidance system that helps a blind user
+find a specific location of interest on an object using CamIO, an
+audio-based AR system. The guidance system uses simple verbal commands
+(``left''/ ``right''/``up''/etc.) to tell the user which direction to
+move the stylus tip. Experiments with four blind participants show that
+the guidance system significantly speeds up the search for target
+hotspots compared with an exhaustive search. A secondary contribution of
+this study is to demonstrate the ability of the CamIO stylus to access
+recessed or otherwise obscured locations on an object that are not
+themselves visible to the camera.
 
-While we applied this guidance approach specifically to CamIO, it could be useful in any AR system, whether the system tracks the user’s hands, handheld stylus or other pointing tool.
-Such audio-based guidance is not only helpful for blind users but is also useful in applications where visibility is limited (either by environmental conditions or because of a visual impairment), or for sighted users who prefer a cross-sensory display (e.g., audio and visual combined).
+While we applied this guidance approach specifically to CamIO, it could
+be useful in any AR system, whether the system tracks the user's hands,
+handheld stylus or other pointing tool. Such audio-based guidance is not
+only helpful for blind users but is also useful in applications where
+visibility is limited (either by environmental conditions or because of
+a visual impairment), or for sighted users who prefer a cross-sensory
+display (e.g., audio and visual combined).
 
 Future work will focus on refining and improving the audio interface.
-One possible way to speed up the guidance process is to add more specific directional feedback, e.g., “far left” would tell the user to move the stylus farther to the left than the “left” instruction.
-The verbal directions could be augmented with 3D spatialized sound as in \cite{mayAuditoryDisplaysFacilitate2019}, which could make the guidance more intuitive.
-Finally, we will continue to test and improve the CamIO system, with an emphasis on refining the spatial accuracy of the tip location estimates using improved computer vision algorithms.
+One possible way to speed up the guidance process is to add more
+specific directional feedback, e.g., ``far left'' would tell the user to
+move the stylus farther to the left than the ``left'' instruction. The
+verbal directions could be augmented with 3D spatialized sound as in
+\cite{mayAuditoryDisplaysFacilitate2019}, which could make the guidance
+more intuitive. Finally, we will continue to test and improve the CamIO
+system, with an emphasis on refining the spatial accuracy of the tip
+location estimates using improved computer vision algorithms.
+%
+% ---- Bibliography ----
+%
+% BibTeX users should specify bibliography style 'splncs04'.
+% References will then be sorted and formatted in the correct style.
+%
+\bibliographystyle{res/templates/splncs04}
+% \bibliography{bibliography}
+\bibliography{res/refs.bib}
+
+\end{document}

### 2020-05-28 R/00-create_data.Rmd
Init

```{r}
conditions <- c('1n', '10g', '2n', '4g', '6n', '7g', '8n', '8g', '9n', '3g', '10n', '5g', '7n', '2g', '5n', '6g', '4n', '9g', '3n', '1g')
#Amy, Khoja, Sarah, Brandon
P1 <- c(120,  20,    120,   53,  120,   41,   120,  11,   120,  13,   120,   14,   120, 14,   120,   16,  120,   56,   120,  31)
P2 <- c(120,  13,    120,   20,  120,   40,   120,  14,   120,  1,   120,    16,   120, 12,   120,   18,  120,   35,   98,   7)
P3 <- c(120,  2,     73,    24,  120,   20,   120,  24,    13,  8,    71,     9,   120, 29,   120,   16,  120,   9,    59,   10)
P4 <- c(82,   14,    120,   15,  70,    28,   120,   9,     4,  1,    120,    9,    14, 15,    47,   15,  120,   9,    118,  6)

data <- data.frame(conditions, P1, P2, P3, P4)
```

```{r}
# Splitting your "condition" variable into two separate factors: hotspot & mode (G vs NG)
data_wide <- data %>% 
  mutate(hotspot = as.factor(str_extract(conditions, "[0-9]+"))) %>% 
  mutate(mode = as.factor(toupper(str_extract(conditions, "[aA-zZ]+")))) %>% 
  #mutate(index = as.factor(seq.int(nrow(.)))) %>%
  select(hotspot, mode, everything(), -conditions)
```

```{r}
# Save filtered data as .csv file
write.csv(data_wide, 'data/camio.csv', row.names = TRUE)
```

### 2020-05-28 R/01-visualizations.Rmd
Init

## Data:
***

Original data:

```{r}
DT::datatable(
  data.wide,
  class = 'cell-border stripe compact',
  filter = 'top',
  options = list(pageLength = 20, autoWidth = TRUE),
  caption = htmltools::tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    glue('Fig {tc<-tc+1}: Filtered data (wide format)'))
)
```

Long format:

```{r}
DT::datatable(
  data.long,
  class = 'cell-border stripe compact',
  filter = 'top',
  options = list(pageLength = 20, autoWidth = TRUE),
  caption = htmltools::tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    glue('Fig {tc<-tc+1}: Data (long format)'))
)
```

## Global summary statistics:
***

```{r}
desc <- data.long %>%
  group_by(mode) %>%
  summarize(
    n = n(),
    mean = mean(time),
    median = median(time),
    sd = sd(time),
    skew = skewness(time),
    kurt = kurtosis(time),
    SW.p = shapiro.test(time)$p.value
  )

DT::datatable(
  desc,
  class = 'cell-border stripe compact',
  filter = 'top',
  options = list(pageLength = 20, autoWidth = TRUE),
  caption = htmltools::tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    glue('Fig {tc<-tc+1}: Summary statistics'))
) %>% 
  formatRound(2:ncol(desc), 5) %>%
  formatStyle(
    'SW.p',
    color = styleInterval(alpha, c('red', 'gray')),
    fontWeight = styleInterval(alpha, c('bold', 'normal')),
  )
```

[!] TODO: better summary table
|   See [here](https://dabblingwithdata.wordpress.com/2018/01/02/my-favourite-r-package-for-summarising-data/)

```{r}
#data.long %>% tbl_summary %>% bold_labels
#summary(arsenal::tableby(mode ~ hotspot, data.long), title = "Arsenal summary")
table1::table1(~time | mode + subject, data = data.long)
```

## Summary statistics per guidance mode:
***

### Distribution of scores per guidance mode :

```{r}
fig <- data.long %>% 
  group_by(mode) %>% 
  mutate(med = median(time)) %>% 
  ungroup() %>%
  ggplot(aes(x=time, fill=mode, color=mode)) +
  geom_histogram(aes(y=..density..), binwidth=15, alpha=0.8, fill="white", boundary = 1) +
  geom_density(alpha=0.6, adjust=6) + 
  theme(legend.position = "none") +
  labs(title= glue('Fig {tc<-tc+1}: Score distributions + median values per type of guidance'), x = "Time to find hotspot") +
  facet_wrap(. ~ mode) +
  geom_vline(aes(xintercept=med, group=mode, color=mode, label="100"), linetype="dashed", size=1) +
  geom_text(aes(x=med - 10, label=med, y=0.055))

ggplotly(fig)
```

### Violin & Box plots per guidance mode:

```{r}
plot_ly(
    data.long,
    x = ~mode,
    y = ~time,
    color = ~mode,
    type = 'violin',
    points = 'all',
    box = list(visible = T),
    meanline = list(visible = T),
    showlegend = FALSE,
    colors = "Set1"
  ) %>%
  plotly::layout(
    xaxis = list(title = "Subjects"),
    yaxis = list(title = "Time", range = c(0,120)),
    title = glue('Fig {tc<-tc+1}: Violin + Box-Plots per guidance mode')
  )
```

### Summary of scores per guidance mode :

```{r}
summary.mode <- data.long %>%
  group_by(mode) %>%
  summarize(
    n = n(),
    median = median(time),
    mean = mean(time),
    skew = skewness(time),
    kurt = kurtosis(time),
    SW.p = shapiro.test(time)$p.value
  )

DT::datatable(
  summary.mode,
  class = 'cell-border stripe compact',
  options = list(autoWidth = TRUE),
  caption = htmltools::tags$caption(style = 'caption-side: bottom; text-align: center;',
                                    glue('Fig {tc<-tc+1}: Summary statistics per guidance mode'))
) %>%
  formatRound(5:ncol(summary.mode), 4) %>%
  formatStyle('SW.p',
              color = styleInterval(alpha, c('red', 'gray')),
              fontWeight = styleInterval(alpha, c('bold', 'normal')))
```

## Summary statistics per subject:
***

### Distribution per subject (guided condition):

```{r}
fig <- data.long %>% 
  filter(mode == "G") %>%
  group_by(subject) %>% 
  mutate(med = median(time)) %>% 
  ungroup() %>%
  ggplot(aes(x=time, fill=subject, color=subject)) +
  geom_histogram(aes(y=..density..), binwidth=5, alpha=0.8, fill="white", boundary = 1, position="identity") +
  geom_density(aes(fill=subject), alpha=0.6, adjust=2) + 
  theme(legend.position = "none") +
  facet_wrap(. ~ subject, ncol=2) +
  geom_vline(aes(xintercept=med, color=subject), linetype="dashed", size=1) +
  geom_text(aes(x=med - 5, label=med, y=-0.01)) +
  labs(title= glue('Fig {tc<-tc+1}: Score distributions & median scores per subject (guided mode)'), x = "Time to find hotspot")

ggplotly(fig)
```

### Violin & Box plots per subject (guided condition):

```{r}
plot_ly(
    data.long %>% filter(mode == "G"),
    x = ~subject,
    y = ~time,
    color = ~subject,
    type = 'violin',
    points = 'all',
    box = list(visible = T),
    meanline = list(visible = T),
    showlegend = FALSE,
    colors = "Set1"
  ) %>%
  plotly::layout(
    xaxis = list(title = "Subjects"),
    yaxis = list(title = "Time", range = c(0,120)),
    title = glue('Fig {tc<-tc+1}: Violin + Box-Plots (guided condition)')
  )
```

### Summary per subject (guided condition):

```{r}
summary.subject <- data.long %>%
  filter(mode == "G") %>%
  group_by(subject) %>%
  summarize(
    n = n(),
    median = median(time),
    mean = mean(time),
    skew = skewness(time),
    kurt = kurtosis(time),
    SW.p = shapiro.test(time)$p.value
  )

DT::datatable(
  summary.subject,
  class = 'cell-border stripe compact',
  options = list(autoWidth = TRUE),
  caption = htmltools::tags$caption(style = 'caption-side: bottom; text-align: center;',
                                    glue('Fig {tc<-tc+1}: Summary statistics per subject'))
) %>%
  formatRound(5:ncol(summary.subject), 4) %>%
  formatStyle('SW.p',
              color = styleInterval(alpha, c('red', 'gray')),
              fontWeight = styleInterval(alpha, c('bold', 'normal')))
```

```{r}
## Violin + Box plots per subject * mode

# violin <- data.long %>%
#   plot_ly(
#     type = 'violin',
#     points = 'all',
#     #color = ~subject,
#     box = list(visible = T),
#     meanline = list(visible = T),
#     showlegend = FALSE,
#     colors = "Set1"
#   )
# 
# violin %<>%
#   add_trace(
#     x = ~subject[data.long$mode == 'G'],
#     y = ~time[data.long$mode == 'G'],
#     legendgroup = 'G',
#     scalegroup = 'G',
#     name = 'G',
#     color = I("blue")
#     # box = list(
#     #   visible = T
#     # ),
#     # meanline = list(
#     #   visible = T
#     # )
#   )
# 
# violin %<>%
#   add_trace(
#     x = ~subject[data.long$mode == 'N'],
#     y = ~time[data.long$mode == 'N'],
#     legendgroup = 'N',
#     scalegroup = 'N',
#     name = 'N',
#     color = I("pink")
#     # box = list(
#     #   visible = T
#     # ),
#     # meanline = list(
#     #   visible = T
#     # ),
#   ) 
# 
# violin %<>%
#   layout(
#     xaxis = list(title = "Subject and Guidance mode"),
#     yaxis = list(title = "Time", range = c(0,130)),
#     title = glue('Figure {tc<-tc+1}: Violin + Box-Plots for both conditions'),
#     violinmode = 'group'
#   )
# 
# violin
```

## Summary statistics per hotspot:
***

### Distribution per hotspot (guided condition):

```{r fig.width=10, fig.height=10}
fig <- data.long %>%
  filter(mode == "G") %>%
  group_by(hotspot) %>% 
  mutate(med = median(time)) %>% 
  ungroup() %>%
  ggplot(aes(x=time, fill=hotspot, color=hotspot)) +
  geom_histogram(aes(y=..density..), binwidth=5, alpha=0.8, fill="white", boundary = 1, position="identity") +
  geom_density(aes(fill=hotspot), alpha=0.6, adjust=3) + 
  theme(legend.position = "none") +
  facet_wrap(. ~ hotspot, ncol=4) +
  geom_vline(aes(xintercept=med, color=hotspot), linetype="dashed", size=1) +
  geom_text(aes(x=med + 10, label=med, y=-0.02)) +
  labs(title = glue('Fig {tc<-tc+1}: Score distributions & median scores per hotspot (guided mode)'),
    x = "Time to find hotspot")

ggplotly(fig)
```

### Violin & Box plots per hotspot (guided mode):

```{r}
plot_ly(
    data.long %>% filter(mode == "G"),
    x = ~hotspot,
    y = ~time,
    color = ~hotspot,
    type = 'violin',
    points = 'all',
    box = list(visible = T),
    meanline = list(visible = T),
    showlegend = FALSE,
    colors = "Set1"
  ) %>%
  layout(
    xaxis = list(title = "Hotspots"),
    yaxis = list(title = "Time", range = c(0,120)),
    title = glue('Fig {tc<-tc+1}: Violin + Box-Plots per hotspot (guided mode)')
  )
```

### Summary per hotspot (guided mode):

```{r}
summary.hotspots <- data.long %>%
  filter(mode == "G") %>%
  group_by(hotspot) %>%
  summarize(
    n = n(),
    median = median(time),
    mean = mean(time),
    skew = skewness(time),
    kurt = kurtosis(time),
    SW.p = shapiro.test(time)$p.value
  )

DT::datatable(
  summary.hotspots,
  class = 'cell-border stripe compact',
  options = list(autoWidth = TRUE),
  caption = htmltools::tags$caption(style = 'caption-side: bottom; text-align: center;',
                                    glue('Fig {tc<-tc+1}: Summary statistics per hotspot (guided mode)'))
) %>%
  formatRound(5:ncol(summary.hotspots), 4) %>%
  formatStyle('SW.p',
              color = styleInterval(alpha, c('red', 'gray')),
              fontWeight = styleInterval(alpha, c('bold', 'normal')))
```

### 2020-05-28 R/02-analysis.Rmd
Init

## Scores as a function of time / trial :
***

### Linear fit on time/trials:

```{r}
data.long.time.G <- data.long %>% 
  group_by(subject, mode) %>% 
  mutate(id = seq_along(hotspot)) %>% 
  ungroup() %>%
  select(id, everything()) %>%
  filter(mode == "G")
```

```{r}
lin.mod.time <- lm(time ~ id, data.long.time.G)
```

```{r}
broom::tidy(lin.mod.time) %>%
  DT::datatable(
  class = 'cell-border stripe compact',
  options = list(autoWidth = TRUE),
  caption = htmltools::tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    glue('Fig {tc<-tc+1}: LM time model summary')
  )
) %>% formatRound(2:5, 4)
```

```{r}
fig <- data.long.time.G %>% 
  ggplot(aes(x = id, y = time, color = subject)) +
  labs(title= glue('Fig {tc<-tc+1}: Score evolution per trial (guided mode)'), x = "Trials") +
  theme(legend.position = "none") +
  facet_wrap(. ~ subject, ncol = 2) +
  geom_point() +
  geom_smooth(
    method = "lm",
    formula = y ~ x,
    #family = gaussian(link = 'log'),
    alpha = .15,
    aes(fill = subject)
  )

ggplotly(fig)
```

#### (TODO) Fit metrics:

[!] TODO: fit per subject

```{r}
broom::glance(lin.mod.time) %>% 
  DT::datatable(
  class = 'cell-border stripe compact',
  options = list(autoWidth = TRUE),
  caption = htmltools::tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    glue('Fig {tc<-tc+1}: Linear model evaluation summary')
  )
) %>% formatRound(1:10, 4)
```

Overall, there's a very bad linear fit on trials/time with current data. We'd need more values see an eventual temporal trend in the data, with the high amount of variations in scores per hotspot & per subject.

### (TODO) Linear fit with autocorrelation:

[!] TODO: see [here](https://www.r-bloggers.com/linear-regression-with-correlated-data/)

```{r}
#lin.mod.time.ar <- nlme::gls(time ~ mode, correlation=nlme::corAR1(form=~1), data = data.long.time.G)
#anova(lin.mod.time, lin.mod.time.ar)
```

### Exponential fit on time/trials:

No convergence with current data

```{r}
# Estimating the initial parameters (using a linear model)
# TODO: check nlstools

##exp.forumla <- alpha * exp(beta * x) + theta

# theta.0 <- min(dt_time$time) * 0.5
# 
# model.0 <- lm(log(time - theta.0) ~ id, data=dt_time)  
# alpha.0 <- exp(coef(model.0)[[1]])
# beta.0 <- coef(model.0)[[2]]
# 
# start <- c(alpha = alpha.0, beta = beta.0, theta = theta.0)
# print(start)
```

```{r}
# fig <- ggplot(data_P1, aes(x = id)) +
#   labs(title= glue('Fig {tc<-tc+1}: Test'), x = "Trials") +
#   geom_point(aes(y = time)) +
#   stat_function(fun = function(x) start[1] * exp(start[2] * x) + start[3], n = 100, color="red")
# 
# ggplotly(fig)
```

```{r}
# exp.model <- nlme::gnls(time ~ alpha * exp(beta * id) + theta, data=data_P1, start=start)
# exp.model
```

### Sigmoid fit on time/trials:

No convergence with current data
```{r}
# sig.mod <- gnls(time ~ SSlogis(id, asym, xmid, scal), data=data_P1, start=list(a=1,b=.5,c=25))
# summary(sig.mod)
```

```{r}
# fig <- ggplot(data_P1, aes(x = id)) +
#   labs(title= glue('Fig {tc<-tc+1}: Sigmoid fit'), x = "Trials") +
#   geom_point(aes(y = time)) +
#   stat_function(fun = function(x) params[1] / (1 + exp(-params[2] * (x - params[3]))), n = 100, color="blue")
# 
# ggplotly(fig)
```

Both sigmoid & exponential fit don't converge with the current data.

## Pairwise differences evaluation:
***

### Pairwise differences summary :

Paired by hotspot

```{r}
pairwise <- data.long %>%
  group_by(subject, hotspot) %>%
  summarize(diff = abs(diff(time))) %>%
  ungroup()
```

```{r}
DT::datatable(
  pairwise,
  class = 'cell-border stripe compact',
  options = list(autoWidth = TRUE),
  caption = htmltools::tags$caption(style = 'caption-side: bottom; text-align: center;',
                                    glue('Fig {tc<-tc+1}: Summary statistics of pairwise differences'))
)
```

### Pairwise differences distribution :

```{r}
pairwise.fig <- pairwise %>%
  ggplot(aes(x=diff)) +
  geom_histogram(aes(y=..density..), binwidth=5, alpha=0.9, color="blue", fill="blue", boundary = 1) +
  geom_density(alpha=0.6, adjust=1.2, color="blue", fill="blue") + 
  labs(title= glue('Fig {tc<-tc+1}: Distribution of pairwise score differences'), x = "Pairwise difference in time to find the hotspot")

ggplotly(pairwise.fig)
```

#### Violin + Box-plot:

```{r}
plot_ly(
    pairwise,
    y = ~diff,
    type = 'violin',
    points = 'all',
    box = list(visible = T),
    meanline = list(visible = T),
    colors = "Set1"
  ) %>%
  layout(
    yaxis = list(title = "Pairwise time difference", range=c(0,120)),
    title = glue('Fig {tc<-tc+1}: Violin + Box-Plots of pairwise differences')
  )
```

### Pairwise differences normality:

#### QQ Plot:

```{r}
ggplotly(
  ggplot(pairwise, aes(sample = diff)) + 
    geom_qq(color="blue") + 
    geom_qq_line() +
    labs(title= glue('Fig {tc<-tc+1}: QQ plot of score diffrences'))
  )
```

#### Summary:

```{r}
moments <- pairwise %>%
  summarize(
    n = n(),
    median = median(diff),
    mean = mean(diff),
    skew = skewness(diff),
    kurt = kurtosis(diff),
    SW.p = shapiro.test(diff)$p.value
  )

DT::datatable(
  moments,
  class = 'cell-border stripe compact',
  options = list(autoWidth = TRUE),
  caption = htmltools::tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    glue('Fig {tc<-tc+1}: Summary statistics of score differences')
  )
) %>%
  formatRound(4:ncol(moments), 4) %>%
  formatStyle('SW.p',
              color = styleInterval(alpha, c('red', 'gray')),
              fontWeight = styleInterval(alpha, c('bold', 'normal')))
```

### Box-Cox transformation:

Box-Cox transform with $\lambda = 6$

```{r}
pairwise.boxcox <- pairwise %>%
  mutate(bc.diff = BoxCox(diff, 6))
```

```{r}
pairwise.boxcox.fig <- pairwise.boxcox %>%
  ggplot(aes(x=bc.diff)) +
  geom_density(alpha=0.8, color="blue", fill="blue", adjust=1.2) +
  labs(title= glue('Fig {tc<-tc+1}: Pairwise score difference distribution (Box-Cox)'), x = "BoxCox(Time to find hotspot)")

ggplotly(pairwise.boxcox.fig)
```

```{r}
ggplotly(
  ggplot(pairwise.boxcox, aes(sample = bc.diff)) + 
    geom_qq(color="blue") + 
    geom_qq_line() +
    labs(title= glue('Fig {tc<-tc+1}: QQ plot of Box-Cox corrected score diffrences'))
  )
```

```{r}
summary.pairwise.boxcox <- pairwise.boxcox %>%
  summarize(
    n = n(),
    skew = skewness(diff),
    kurt = kurtosis(diff),
    SW.p = shapiro.test(diff)$p.value,
    skew.boxcox = skewness(bc.diff),
    kurt.boxcox = kurtosis(bc.diff),
    SW.p.boxcox = shapiro.test(bc.diff)$p.value
  )

DT::datatable(
  summary.pairwise.boxcox,
  class = 'cell-border stripe compact',
  options = list(autoWidth = TRUE),
  caption = htmltools::tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    glue('Fig {tc<-tc+1}: Summary statistics of Box-Cox transformed score differences')
  )
) %>%
  formatRound(2:ncol(summary.pairwise.boxcox), 4) %>%
  formatStyle(c('SW.p', 'SW.p.boxcox'),
              color = styleInterval(alpha, c('red', 'gray')),
              fontWeight = styleInterval(alpha, c('bold', 'normal')))
```

The Box-Cox transform seems to make the date more normal --> to investigate ?

## Non-parametric comparaisons:
***

### Wilcoxon Signed-Rank Test:

```{r}
data.wide.GN <- data.long %>% reshape2::dcast(subject+hotspot~mode,  value.var="time")
#print(data.wide.GN)
```

```{r}
wilcoxon.res <- wilcox.test(data.wide.GN$N, data.wide.GN$G, alternative="two.sided", paired=TRUE, conf.int=TRUE)

summary.wilcoxon <- wilcoxon.res %>% tidy

DT::datatable(
  summary.wilcoxon,
  class = 'cell-border stripe compact',
  options = list(autoWidth = TRUE),
  caption = htmltools::tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    glue('Fig {tc<-tc+1}: Signed-Rank Wilcoxon Test summary')
  )
) %>% formatRound(1:5, 10)
```

#### Effect size:

```{r}
wilcoxon.es <- wilcoxon.res$statistic / sqrt(nrow(data.wide.GN))
#wilcoxon.es2 <- data.long %>% rstatix::wilcox_effsize(time ~ mode, paired=TRUE)
print(wilcoxon.es)
```

Effect-size for Wilcoxon Signed-Rank Test is `r wilcoxon.es`

### (TODO) Resampling test:

[!] TODO

## Parametric models :
***

### LM model:

I'm only fitting a linear model for experimenting: with the current data, almost every assumption of Linear Models is violated.

```{r results=F}
lin.mod <- lm(time ~ mode, data.long)
```

#### Heteroscedasticity:

```{r}
par(mfrow=c(2,2))
plot(lin.mod, 1:2)
```

* Residual vs Fitted: The difference in spread in both conditions points towards heteroscedasticity, so that linear model doesn't hold.
* QQ plot: data is also obviously not normal.

##### Heteroscedasticity test :

```{r}
lmtest::bptest(lin.mod) %>% tidy
```

No surprise on the heteroscedasticity with the right-censoring of the measures.

[!] TODO: try to fix it ?  
- Data transform (Box-Cox, ...) ?   
- GLS Regression ?  
- Truncated regression ?

#### Predicted values:

```{r}
ggpredict(lin.mod, terms = c("mode")) %>% plot(add.data = TRUE)
```

#### Model fit:

```{r}
broom::glance(lin.mod)
```
Significant effect of the guidance mode (but we can't really use that test/model)

AIC of ~747

#### Effects / contrasts :

```{r}
#broom::tidy(car::Anova(lin.mod))
```

```{r}
broom::tidy(lin.mod) %>%
  DT::datatable(
  class = 'cell-border stripe compact',
  options = list(autoWidth = TRUE),
  caption = htmltools::tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    glue('Fig {tc<-tc+1}: LM model summary')
  )
) %>% formatRound(2:5, 4)
```

```{r}
emm1 = emmeans(lin.mod, specs = pairwise ~ mode, adjust = "none", type = "response")

emm1$contrasts %>% 
  summary(infer = TRUE) %>% 
  as.data.frame %>%
  DT::datatable(
  class = 'cell-border stripe compact',
  options = list(autoWidth = TRUE),
  caption = htmltools::tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    glue('Fig {tc<-tc+1}: LM contrast summary')
  )
) %>% formatRound(c(3,5,6,7,8), 4)
```

[!] TODO: bootstrap CI measures ?

```{r}
plot(emm1) + 
  theme_bw() + 
  labs(x = "Estimated time (mean, 95CI)", y = "Condition")
```

### (TODO) GLM model:

[!] TODO: find the optimal familly and link function for the data.

```{r}
glin.mod <- glm(time ~ mode, data.long, family=Gamma(link="log"))
```

```{r}
par(mfrow=c(2,2))
plot(glin.mod, 1:2)
```

#### Model fit:

```{r}
broom::glance(glin.mod) %>% 
  DT::datatable(
  class = 'cell-border stripe compact',
  options = list(autoWidth = TRUE),
  caption = htmltools::tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    glue('Fig {tc<-tc+1}: GLM evaluation summary')
  )
) %>% formatRound(1:10, 4)
```

### (TODO) GLMM model:

[!] TODO

```{r}
data.long.time2 <- data.long %>% 
  group_by(subject, mode) %>% 
  mutate(id = seq_along(hotspot)) %>% 
  ungroup() %>%
  select(id, everything()) %T>% print
```

```{r}
glmm1 <- lmer(time ~mode + (1|id/mode), data=data.long.time2)
glmm1
```

Checking model assumptions (normality of residuals = homeoscedasticity along predicted values)
```{r}
par(mfrow=c(2,2))
plot(glmm1)
```

### Model comparison:

```{r}
t1 <- tbl_regression(lin.mod)
t2 <- tbl_regression(glin.mod, exponentiate = TRUE)

tbl_merge(
    tbls = list(t1, t2),
    tab_spanner = c("**Linear**", "**Generalized (Gaussian, Log link)**")
  ) %>%
  bold_labels()
```

[!] TODO: model fit comparison

```{r}
# adj_R2 <-rbind("exp.model" = summary(exp.model)$adj.r.squared,
#                "sig.model" = summary(sig.model)$adj.r.squared,
#                "lin.model" = summary(lin.model)$adj.r.squared)
# 
# colnames(adj_R2) <- "adj_R2"
# adj_R2
```

### 2020-05-28 R/Index.Rmd
Init

---
title: "CamIO data analysis"
author: "RIVIERE Marc-Aurèle"
output:
  prettydoc::html_pretty:
    highlight: github
    math: mathjax
    code_folding: hide
    number_sections: yes
    theme: cayman
    toc: yes
  html_notebook:
    code_folding: hide
    highlight: tango
    number_sections: yes
    theme: spacelab
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  html_document:
    theme: sandstone
    highlight: tango
    df_print: paged
    toc: yes
always_allow_html: yes
---

# Setup:

```{r setup, include=FALSE, cache=FALSE}

if (!require(pacman)) install.packages("pacman")

pkg_list <- c("glue", "broom", "magrittr", "prettydoc", "DT", "reshape2", "plotly", "moments", "emmeans", "gtsummary", "lmtest", "table1", "rstatix", "lme4", "ggeffects", "tidyverse", "conflicted", "forecast")

pacman::p_load(char = pkg_list) #update=TRUE
pacman::p_load_gh("rstudio/gt")
#knitr::write_bib(c(.packages(), pkg_list), 'bib/packages.bib')

knitr::opts_chunk$set(
  fig.align="center",
  #fig.pos = "!H",
  #out.extra = "",
  tidy = TRUE,
  warning = FALSE,
  echo = FALSE,
  message = FALSE,
  error=FALSE,
  cache = FALSE
)

options(scipen=999, digits=5)

tc <- 0
alpha <- 0.05
set.seed(42)
```

```{r}
sessionInfo()

conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("layout", "plotly")

# includes:
#   in_header: header.html
#   after_body: sig.html
```

# Data preparation:
***
***

```{r}
data.wide <- readr::read_csv("../data/camio.csv") %>% dplyr::select(-X1) %T>% print
```

```{r results=FALSE}
## With reshape2
data.long <- data.wide %>% reshape2::melt(measure.vars=c("P1", "P2", "P3", "P4"), variable.name = "subject", value.name="time") %T>% print

## (old) With tidyr 0.8.5
#data.long <- data.wide %>% gather(key="subject", value="time", "P1":"P4") %T>% print
#data.long <- data.wide %>% gather(key="subject", value="time", c("P1", "P2", "P3", "P4")) %T>% print

## With tidyr 1.0.0
#data.long <- data.wide %>% pivot_longer(cols = "P1":"P4", names_to="subject", values_to = "time") %T>% print
```

# Summary / descriptive statistics:
***
***

```{r child = '01-visualizations.Rmd', echo=TRUE}
```

# Analysis:
***
***

```{r child = '02-analysis.Rmd', echo=TRUE}
```

### 2020-05-29 Manuscript.Rmd
Update stats

@@
 output:
+  word_document:
+    reference_docx: res/templates/template.docx
+    toc: no
   pdf_document:
-    template: res/templates/markdown.tex
+    template: res/templates/template.tex
     toc: no

### 2020-05-29 R/01-visualizations.Rmd
Update stats

@@
 
-```{r}
+```{r warnings=F}
 desc <- data.long %>%
@@
     n = n(),
+    median = median(time),
     mean = mean(time),
-    median = median(time),
     sd = sd(time),
@@
   class = 'cell-border stripe compact',
-  filter = 'top',
-  options = list(pageLength = 20, autoWidth = TRUE),
   caption = htmltools::tags$caption(
@@
 ) %>% 
-  formatRound(2:ncol(desc), 5) %>%
+  formatRound(5:ncol(desc), 4) %>%
   formatStyle(
@@
   facet_wrap(. ~ mode) +
-  geom_vline(aes(xintercept=med, group=mode, color=mode, label="100"), linetype="dashed", size=1) +
+  geom_vline(aes(xintercept=med, group=mode, color=mode), linetype="dashed", size=1) +
   geom_text(aes(x=med - 10, label=med, y=0.055))
@@
   class = 'cell-border stripe compact',
-  options = list(autoWidth = TRUE),
   caption = htmltools::tags$caption(style = 'caption-side: bottom; text-align: center;',
@@
   formatStyle('SW.p',
-              color = styleInterval(alpha, c('red', 'gray')),
+              color = styleInterval(alpha, c('red', 'black')),
               fontWeight = styleInterval(alpha, c('bold', 'normal')))

### 2020-05-29 R/02-analysis.Rmd
Update stats

@@
   select(id, everything()) %>%
-  filter(mode == "G")
+  filter(mode == "G") %T>% print
 ```
@@
 
-[!] TODO: fit per subject
+[!] TODO: fit per subject --> random effect on subject ?
 
@@
   summarize(diff = abs(diff(time))) %>%
-  ungroup()
+  ungroup() %T>% print
 ```
@@
 ```{r}
-data.wide.GN <- data.long %>% reshape2::dcast(subject+hotspot~mode,  value.var="time")
+data.wide.GN <- data.long %>% reshape2::dcast(subject+hotspot~mode,  value.var="time") %T>% print
 #print(data.wide.GN)
@@
   )
-) %>% formatRound(1:5, 10)
+) %>% formatRound(1:5, 2)
 ```

### 2020-05-29 R/Index.Rmd
Update stats

@@
 
-pkg_list <- c("glue", "broom", "magrittr", "prettydoc", "DT", "reshape2", "plotly", "moments", "emmeans", "gtsummary", "lmtest", "table1", "rstatix", "lme4", "ggeffects", "tidyverse", "conflicted", "forecast")
+pkg_list <- c("tidyverse", "glue", "broom", "magrittr", "prettydoc", "DT", "reshape2", "plotly", "moments", "emmeans", "gtsummary", "lmtest", "table1", "rstatix", "lme4", "ggeffects", "conflicted", "forecast")
 
 pacman::p_load(char = pkg_list) #update=TRUE
-pacman::p_load_gh("rstudio/gt")
+pacman::p_load_gh(c("rstudio/gt", "tidyverse/dplyr"))
 #knitr::write_bib(c(.packages(), pkg_list), 'bib/packages.bib')
@@
   #out.extra = "",
-  tidy = TRUE,
+  #tidy = TRUE,
   warning = FALSE,
@@
 set.seed(42)
+
+initial.data <- "../data/initial.csv"
 ```
@@
 ```{r}
-data.wide <- readr::read_csv("../data/camio.csv") %>% dplyr::select(-X1) %T>% print
+data.wide <- readr::read_csv(initial.data) %>% 
+  dplyr::select(-X1) %>% 
+  mutate(across(c(hotspot, mode), as.factor)) %T>% print
 ```
@@
 ```{r results=FALSE}
-## With reshape2
-data.long <- data.wide %>% reshape2::melt(measure.vars=c("P1", "P2", "P3", "P4"), variable.name = "subject", value.name="time") %T>% print
+### With reshape2
+data.long <- data.wide %>% 
+  reshape2::melt(measure.vars=c("P1", "P2", "P3", "P4"), variable.name = "subject", value.name="time") %>% 
+  group_by(subject) %>% mutate(id = seq_along(hotspot), .before = hotspot) %>% ungroup() %T>% print
 
-## (old) With tidyr 0.8.5
+### (old) With tidyr 0.8.5
 #data.long <- data.wide %>% gather(key="subject", value="time", "P1":"P4") %T>% print
@@
 
-## With tidyr 1.0.0
-#data.long <- data.wide %>% pivot_longer(cols = "P1":"P4", names_to="subject", values_to = "time") %T>% print
+### With tidyverse >= 1.0.0 && dplyr >= 0.8.99
+# data.long <- data.wide %>% 
+#   mutate(id = seq_along(hotspot), .before = hotspot) %>% 
+#   pivot_longer(cols = "P1":"P4", names_to="subject", names_transform = list(subject = as.factor), values_to = "time") %>%
+#   group_by(subject) %>% arrange(id, .by_group=T) %>% ungroup() %T>% print
 ```

### 2020-05-30 Manuscript.Rmd
Updated template with \thanks

--- 2020-05-28 Manuscript.Rmd
@@
 authorrunning: J. Couhlan et al.
+thanks: Merci !
 output:
   pdf_document:
-    template: res/templates/markdown.tex
+    template: res/templates/template.tex
     toc: no
     keep_tex: yes
+  word_document:
+    reference_docx: res/templates/template.docx
+    toc: no
 institutes:

### 2020-05-30 R/01-visualizations.Rmd
Update vizs

@@
   class = 'cell-border stripe compact',
-  filter = 'top',
   options = list(pageLength = 20, autoWidth = TRUE),
@@
 
-### Violin & Box plots per guidance mode:
+### Box-Plots per subject & guidance mode:
 
@@
     data.long,
-    x = ~mode,
+    x = ~subject:mode,
     y = ~time,
-    color = ~mode,
-    type = 'violin',
-    points = 'all',
-    box = list(visible = T),
-    meanline = list(visible = T),
+    color = ~subject:mode,
+    type = 'box',
     showlegend = FALSE,
@@
   )
+```
+
+### Lineplot of averages between guidance mode :
+
+```{r}
+dat <- data.long %>% Rmisc::summarySE(measurevar="time", groupvars=c("mode", "subject")) %>% print
+#dat2 <- data.long %>% mutate(se = se(time))
+pd <- position_dodge(0.15)
+
+fig <- dat %>%
+  ggplot(aes(x = mode, y = time, color = subject)) + #group=1
+  geom_line(position = pd) +
+  geom_errorbar(aes(ymin = time - se, ymax = time + se), width = .1, position = pd) +
+  geom_point(shape=21, size=2, fill="white", position = pd)
+
+ggplotly(fig)
 ```

### 2020-05-30 R/Index.Rmd
Update vizs

@@
 
-pkg_list <- c("tidyverse", "glue", "broom", "magrittr", "prettydoc", "DT", "reshape2", "plotly", "moments", "emmeans", "gtsummary", "lmtest", "table1", "rstatix", "lme4", "ggeffects", "conflicted", "forecast")
+pkg_list <- c("tidyverse", "glue", "broom", "magrittr", "prettydoc", "DT", "reshape2", "plotly", "moments", "emmeans", "gtsummary", "lmtest", "table1", "rstatix", "lme4", "ggeffects", "conflicted", "forecast", "Rmisc")
 

### 2020-06-01 manuscript.tex
Update structure: Overleaf first

--- 2020-05-28 Manuscript.tex
@@
-% This is template.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-
 \documentclass[runningheads]{res/templates/llncs}
-
+%
 \usepackage{graphicx}
@@
 \usepackage{hyperref}
-\usepackage{etoolbox}
-
-% Include packages from YAML header
 
@@
 
-\title{An Audio-Based 3D Spatial Guidance AR System for Blind Users}
+\title{An Audio-Based 3D Spatial Guidance AR System for Blind Users\thanks{Thank you for ...}}
+\titlerunning{Audio-Based 3D Spatial Guidance}
 
-\titlerunning{Audio-Based 3D Spatial Guidance}
+\author{James Coughlan\inst{1}\orcidID{0000-0003-2775-4083} \and
+Brandon Biggs\inst{1} \and
+Marc-Aurèle Rivière\inst{2}\orcidID{0000-0002-5108-3382} \and
+Huiying Shen\inst{1}}
+
+\authorrunning{J. Coughlan et al.}
+
+\institute{The Smith-Kettlewell Eye Research Institute, San Francisco, CA \\
+\email{\{coughlan, brandon.biggs, hshen\}@ski.org} \and
+LITIS, University of Rouen, Normandy, France \\
+\email{marc-aurele.riviere@univ-rouen.fr}}
 %
-\newbool{flag}
-\boolfalse{flag}
-%
-\author{
-\ifbool{flag}{\and}{\booltrue{flag}}James Coughlan\inst{1}\orcidID{0000-1111-2222-3333}
-\ifbool{flag}{\and}{\booltrue{flag}}Brandon Biggs\inst{1}\orcidID{0000-1111-2222-3333}
-\ifbool{flag}{\and}{\booltrue{flag}}Marc-Aurèle Rivière\inst{2}\orcidID{0000-0002-5108-3382}
-\ifbool{flag}{\and}{\booltrue{flag}}Huiying Shen\inst{1}\orcidID{0000-1111-2222-3333}
-}
-%
-\authorrunning{J. Couhlan et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\boolfalse{flag}
-%
-\institute{
-\ifbool{flag}{\and}{\booltrue{flag}}
- The Smith-Kettlewell Eye Research Institute, San Francisco, CA \\
- \email{\href{mailto:coughlan@ski.org}{\nolinkurl{coughlan@ski.org}},
-\href{mailto:brandon.biggs@ski.org}{\nolinkurl{brandon.biggs@ski.org}},
-\href{mailto:hshen@ski.org}{\nolinkurl{hshen@ski.org}}} 
-\ifbool{flag}{\and}{\booltrue{flag}}
- LITIS, University of Rouen, Normandy, France \\
- \email{\href{mailto:marc-aurele.riviere@univ-rouen.fr}{\nolinkurl{marc-aurele.riviere@univ-rouen.fr}}} 
-}
-
-%
-\maketitle              % typeset the header of the contribution
+\maketitle %
 %
@@
 for helping users find locations of interest.
-\\
-\boolfalse{flag}
-\keywords{
-\ifbool{flag}{\and}{\booltrue{flag}}
-Assisitive Devices
-\ifbool{flag}{\and}{\booltrue{flag}}
-Augmented Reality
-\ifbool{flag}{\and}{\booltrue{flag}}
-Auditory Substitution
-\ifbool{flag}{\and}{\booltrue{flag}}
-Visual Impairment
-}
+
+\keywords{Assisitive Devices
+\and Augmented Reality
+\and Auditory Substitution
+\and Visual Impairment}
 \end{abstract}
 %
-%
+
 \hypertarget{sota}{%
@@
 \begin{figure}
-\includegraphics[width=0.55\linewidth]{res/img/fig1a} \includegraphics[width=0.45\linewidth]{res/img/fig1b} \caption{(left) Blind participant uses CamIO to explore a toy fire truck.
+\includegraphics[width=0.55\linewidth]{res/img/fig1a}
+\includegraphics[width=0.45\linewidth]{res/img/fig1b} 
+\caption{(left) Blind participant uses CamIO to explore a toy fire truck.
 The participant holds the CamIO stylus (a barcode pattern printed on a square wooden dowel with a pointy tip) to point to possible hotspots on the truck.
@@
 (right) Stylus points to steering wheel in the truck cabin.
-Though the steering wheel is in a recessed location that is not visible to CamIO, the stylus tip location can still be estimated.}\label{fig:exp}
+Though the steering wheel is in a recessed location that is not visible to CamIO, the stylus tip location can still be estimated.}
+\label{fig:exp}
 \end{figure}
@@
 location estimates using improved computer vision algorithms.
-%
+
 % ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
 \bibliographystyle{res/templates/splncs04}
-% \bibliography{bibliography}
 \bibliography{res/refs.bib}

### 2020-06-01 res/manuscript.Rmd
Update structure: Overleaf first

--- 2020-05-28 Manuscript.Rmd
@@
 authorrunning: J. Couhlan et al.
+thanks: Merci !
 output:
   pdf_document:
-    template: res/templates/markdown.tex
+    template: templates/template.tex
     toc: no
     keep_tex: yes
+  word_document:
+    reference_docx: templates/template.docx
+    toc: no
 institutes:
@@
 - name: James Coughlan
-  orcid: 0000-1111-2222-3333
+  orcid: 0000-0003-2775-4083
   inst: 1
 - name: Brandon Biggs
-  orcid: 0000-1111-2222-3333
   inst: 1
@@
 - name: Huiying Shen
-  orcid: 0000-1111-2222-3333
   inst: 1
-ressources: res/
-templates: res/templates/
-bibliography: res/refs.bib
+templates: templates/
+bibliography: refs.bib
 ---
@@
 
-knitr::include_graphics("res/img/fig1a.jpg")
-knitr::include_graphics("res/img/fig1b.jpg")
+knitr::include_graphics("img/fig1a.jpg")
+knitr::include_graphics("img/fig1b.jpg")
 ```
@@
 \begin{figure}
-\includegraphics[width=5.5cm, keepaspectratio,]{res/img/fig2a.jpg}
-\includegraphics[width=6.5cm, keepaspectratio,]{res/img/fig2b.jpg}
+\includegraphics[width=5.5cm, keepaspectratio,]{img/fig2a.jpg}
+\includegraphics[width=6.5cm, keepaspectratio,]{img/fig2b.jpg}
 \caption{(left) Plot of time (x axis) to find target hotspot location for G (‘o’ symbol) vs.

### 2020-06-09 manuscript.tex
Update tex

@@
 
-\title{An Audio-Based 3D Spatial Guidance AR System for Blind Users\thanks{Thank you for ...}}
+\title{An Audio-Based 3D Spatial Guidance AR System for Blind Users\thanks{JMC, BB and HS were supported by NIH grant 5R01EY025332 and NIDILRR grant 90RE5024-01-00.
+We thank Dr.
+Roberto Manduchi, Dr.
+Natela Shanidze, Dr.
+Santani Teng and Dr.
+Ali Cheraghi for helpful suggestions about the experiments.}}
+
 \titlerunning{Audio-Based 3D Spatial Guidance}
 
-\author{James Coughlan\inst{1}\orcidID{0000-0003-2775-4083} \and
-Brandon Biggs\inst{1} \and
+\author{James M. Coughlan\inst{1}\orcidID{0000-0003-2775-4083} \and
+Brandon Biggs\inst{1}\orcidID{0000-0001-7380-1077} \and
 Marc-Aurèle Rivière\inst{2}\orcidID{0000-0002-5108-3382} \and
-Huiying Shen\inst{1}}
+Huiying Shen\inst{1}\orcidID{0000-0002-2195-7085}}
 
-\authorrunning{J. Coughlan et al.}
+\authorrunning{J. M. Coughlan et al.}
 
@@
 
-\keywords{Assisitive Devices
+\keywords{Assistive Devices
+\and Accessibility
 \and Augmented Reality
 \and Auditory Substitution
-\and Visual Impairment}
+\and Visual Impairment
+\and Blindness
+\and Low Vision}
 \end{abstract}
@@
 interest is mounted on a board, which is a printed barcode pattern
-(\ref{fig:exp}) that allows CamIO to estimate the object's pose (3D
+(Fig. \ref{fig:exp}) that allows CamIO to estimate the object's pose (3D
 translation and 3D orientation) relative to the camera. The iPhone is
 typically mounted on a tripod or a gooseneck cellphone holder
-\ref{fig:exp} to obtain a clear view of the board and object, but in
+Fig. ref{fig:exp} to obtain a clear view of the board and object, but in
 some cases \cite{coughlanTowardsAccessibleAudioLabeling2020} users may
 hold the iPhone by hand. The user points the tip of a passive wooden
-stylus \ref{fig:exp}, covered with another barcode pattern, to different
+stylus Fig. \ref{fig:exp}, covered with another barcode pattern, to different
 locations on an object. CamIO estimates the stylus tip's 3D location
@@
 A second stylus (not shown, identical in form but distinct from the one
-shown in \ref{fig:exp}) is used to create new hotspots, and allows the
+shown in Fig. \ref{fig:exp}) is used to create new hotspots, and allows the
 user to make an audio recording to associate with each new hotspot. Our
@@
 We define the guidance interface as follows. The board (barcode pattern
-beneath the truck, see \ref{fig:exp}) defines an xyz coordinate frame,
+beneath the truck, see Fig. \ref{fig:exp}) defines an xyz coordinate frame,
 with \(+x\) = right, \(-x\) = left, \(+y\) = forward (away from the
@@
 guidance feedback and to determine whether the feedback speeds up the
-target search. The experiment \ref{fig:exp} compared how much time a
-blind participant needed to find hotspots (points of interest) on a 3D
+target search. The experiment Fig. \ref{fig:exp} compared how much time a blind participant needed to find hotspots (points of interest) on a 3D
 object with spatial guidance vs.~without it. The object was a large toy
@@
 hotspots were in recessed locations on the truck, such as the shift
-stick and steering wheel (\ref{fig:exp}) in the front cabin, or on the
+stick and steering wheel (Fig. \ref{fig:exp}) in the front cabin, or on the
 side of the truck facing away from the participant; these hotspots were
@@
 (guidance provided) and NG (no guidance). The participant was given up
-to 120 sec.~to find each hotspot, after which a time-out was declared.
+to 120 sec to find each hotspot, after which a time-out was declared.
 In order to minimize the influence of any possible learning effects,
@@
 from moving in the suggested guidance direction. For instance, if in
-\ref{fig:exp} the system directs the participant to move ``forward''
+Fig. \ref{fig:exp} the system directs the participant to move ``forward''
 towards the left side mirror, the stylus will soon hit the cabin
@@
 included 10 trials of both conditions (G and NG), for a total of 20
-trials per participant. After completing the 20 search trials, we
+trials per participant.
+We  placed a visual barrier between the participant and the object while the experimenter defined the hotspot in each trial to prevent participants from using any residual vision to see the location of the hotspot.
+After completing the 20 search trials, we
 administered a System Usability Scale (SUS) \cite{brooke1996sus} to
@@
 guidance feedback and to determine whether the feedback speeds up the
-target search. The experiment (\ref{fig:exp}) compared how much time a
+target search. The experiment (Fig. \ref{fig:exp}) compared how much time a
 blind participant needed to find hotspots (points of interest) on a 3D

### 2020-06-09 R/01-visualizations.Rmd
Update figures

@@
 ```{r warnings=F}
+library(pillar)
+options(pillar.sigfig=3)
+
 desc <- data.long %>%
@@
     SW.p = shapiro.test(time)$p.value
-  )
+  ) %>% mutate(across(where(is.character), as.numeric))
+
+td <- tabl(desc)
 
@@
 ) %>% 
-  formatRound(5:ncol(desc), 4) %>%
-  formatStyle(
-    'SW.p',
-    color = styleInterval(alpha, c('red', 'gray')),
-    fontWeight = styleInterval(alpha, c('bold', 'normal')),
-  )
+  formatRound(4:ncol(desc)-1, 2) %>%
+  formatRound(ncol(desc), 4)
+  # formatStyle(
+  #   'SW.p',
+  #   color = styleInterval(alpha, c('red', 'gray')),
+  #   fontWeight = styleInterval(alpha, c('bold', 'normal')),
+  # )
+
+# png("../res/img/tab1.png", width=480, height=480, bg = "white")
+# grid.table(desc)
+# dev.off()
 ```
@@
   ggplot(aes(x=time, fill=mode, color=mode)) +
-  geom_histogram(aes(y=..density..), binwidth=15, alpha=0.8, fill="white", boundary = 1) +
+  geom_histogram(binwidth=12, fill="white", boundary = 1) + #aes(y=..density..)
   geom_density(alpha=0.6, adjust=6) + 
   theme(legend.position = "none") +
-  labs(title= glue('Fig {tc<-tc+1}: Score distributions + median values per type of guidance'), x = "Time to find hotspot") +
+  labs(x = "Time to find hotspot", y="Count") +
   facet_wrap(. ~ mode) +
   geom_vline(aes(xintercept=med, group=mode, color=mode), linetype="dashed", size=1) +
-  geom_text(aes(x=med - 10, label=med, y=0.055))
+  geom_text(aes(x=med - 10, label=med, y=-1.5))
 
+ggsave(plot=fig, filename="../res/img/histNG.png", device="png", dpi=320, scale=1)
 ggplotly(fig)
@@
   geom_errorbar(aes(ymin = time - se, ymax = time + se), width = .1, position = pd) +
-  geom_point(shape=21, size=2, fill="white", position = pd)
+  geom_point(shape=21, size=2, fill="white", position = pd) +
+  labs(x="Guidance mode", y="Time to find hotspot")
 
 ggplotly(fig)
+ggsave(plot=fig, filename="../res/img/lines.png", device="png", dpi=320, scale=1)
 ```

### 2020-06-09 R/Index.Rmd
Update figures

--- 2020-05-29 R/Index.Rmd
@@
 
-```{r setup, include=FALSE, cache=FALSE}
+```{r include=FALSE}
 
+if (!require(knitr)) install.packages("knitr")
 if (!require(pacman)) install.packages("pacman")
-
-pkg_list <- c("tidyverse", "glue", "broom", "magrittr", "prettydoc", "DT", "reshape2", "plotly", "moments", "emmeans", "gtsummary", "lmtest", "table1", "rstatix", "lme4", "ggeffects", "conflicted", "forecast")
+pkg_list <- c("tidyverse", "glue", "broom", "magrittr", "prettydoc", "DT", "reshape2", "plotly", "moments", "emmeans", "gtsummary", "lmtest", "table1", "rstatix", "lme4", "ggeffects", "conflicted", "forecast", "Rmisc", "gridExtra", "kableExtra", "magick", "webshot")
 
 pacman::p_load(char = pkg_list) #update=TRUE
-pacman::p_load_gh(c("rstudio/gt", "tidyverse/dplyr"))
+#pacman::p_load_gh(c("rstudio/gt", "tidyverse/dplyr"))
 #knitr::write_bib(c(.packages(), pkg_list), 'bib/packages.bib')
@@
 
-options(scipen=999, digits=5)
+options(scipen=999, digits=4)
 
@@
 initial.data <- "../data/initial.csv"
+
+conflict_prefer("mutate", "dplyr")
+conflict_prefer("select", "dplyr")
+conflict_prefer("filter", "dplyr")
+conflict_prefer("summarize", "dplyr")
+conflict_prefer("layout", "plotly")
+
+tabl <- function(data, title="Table caption") {
+  kable(data, col.names=gsub("[.]", " ", names(data)), booktabs=T, align="c", digits=3, format.args=list(big.mark=",", scientific=FALSE), escape=F, caption=title, linesep=c("")) %>% 
+    kable_styling(font_size = 10, latex_options = c("striped", "scale_down", "hold_position"), position = "left") %>%
+    row_spec(0, background = "#d4dbde", bold = T, color = "#2b4894")
+}
+options(knitr.table.format = ifelse(knitr::is_latex_output(), "latex", "pandoc"), knitr.kable.NA = "")
 ```
@@
 sessionInfo()
-
-conflict_prefer("select", "dplyr")
-conflict_prefer("filter", "dplyr")
-conflict_prefer("layout", "plotly")
-
-# includes:
-#   in_header: header.html
-#   after_body: sig.html
 ```

### 2020-06-10 R/01-visualizations.Rmd
Update analysis & figs

@@
   ggplot(aes(x=time, fill=mode, color=mode)) +
-  geom_histogram(binwidth=12, fill="white", boundary = 1) + #aes(y=..density..)
+  geom_histogram(binwidth=15, fill="white", boundary = 1) + #aes(y=..density..)
   geom_density(alpha=0.6, adjust=6) + 
   theme(legend.position = "none") +
+  #theme_bw() +
   labs(x = "Time to find hotspot", y="Count") +
+  coord_fixed(ratio=5) +
   facet_wrap(. ~ mode) +

### 2020-06-10 R/02-analysis.Rmd
Update analysis & figs

@@
 ```{r}
-wilcoxon.res <- wilcox.test(data.wide.GN$N, data.wide.GN$G, alternative="two.sided", paired=TRUE, conf.int=TRUE)
+wilcoxon.res <- wilcox.test(data.wide.GN$N, data.wide.GN$G, paired=TRUE, conf.int=TRUE, correct = FALSE)
 
-summary.wilcoxon <- wilcoxon.res %>% tidy
+wilcoxonZ(
+  data.wide.GN$N,
+  data.wide.GN$G,
+  mu = 0,
+  paired = TRUE,
+  exact = FALSE,
+  correct = FALSE,
+  digits = 3
+)
 
-DT::datatable(
-  summary.wilcoxon,
-  class = 'cell-border stripe compact',
-  options = list(autoWidth = TRUE),
-  caption = htmltools::tags$caption(
-    style = 'caption-side: bottom; text-align: center;',
-    glue('Fig {tc<-tc+1}: Signed-Rank Wilcoxon Test summary')
-  )
-) %>% formatRound(1:5, 2)
+wilcoxon.res %>% tidy
+```
+```{r}
+median(data.wide.GN$G)
+median(data.wide.GN$N)
 ```
@@
 ```{r}
-glmm1 <- lmer(time ~mode + (1|id/mode), data=data.long.time2)
+glmm1 <- lmer(time ~mode + (1|subject), data=data.long.time2)
 glmm1

### 2020-06-10 R/ICCHP20.Rmd
Update analysis & figs

---
title: "R Notebook"
output: html_notebook
---

```{r include=FALSE}

if (!require(knitr)) install.packages("knitr")
if (!require(pacman)) install.packages("pacman")
pkg_list <- c("tidyverse", "glue", "magrittr", "plotly", "moments", "emmeans", "conflicted", "gridExtra", "kableExtra", "rcompanion", "broom", "rstatix", "ggpubr", "Rmisc")

pacman::p_load(char = pkg_list) #update=TRUE

knitr::opts_chunk$set(
  warning = FALSE,
  echo = FALSE,
  message = FALSE,
  error=FALSE,
  cache = FALSE
)

options(scipen=999, digits=4)

alpha <- 0.05
set.seed(42)

initial.data <- "../data/initial.csv"

conflict_prefer("mutate", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("summarize", "dplyr")
conflict_prefer("layout", "plotly")

tabl <- function(data, title="Table caption") {
  kable(data, col.names=gsub("[.]", " ", names(data)), booktabs=T, align="c", digits=3, format.args=list(big.mark=",", scientific=FALSE), escape=F, caption=title, linesep=c("")) %>% 
    kable_styling(font_size = 10, latex_options = c("striped", "scale_down", "hold_position"), position = "left") %>%
    row_spec(0, background = "#d4dbde", bold = T, color = "#2b4894")
}
options(knitr.table.format = ifelse(knitr::is_latex_output(), "latex", "pandoc"), knitr.kable.NA = "")
```

```{r results='hide'}
data.wide <- read_csv(initial.data) %>% 
  select(-X1) %>% 
  mutate(mode = ifelse(mode == "N", "NG", "G")) %>%
  mutate(across(c(hotspot, mode), as.factor))

data.long <- data.wide %>%
  mutate(id = seq_along(hotspot), .before = hotspot) %>%
  pivot_longer(cols = "P1":"P4", names_to="subject", names_transform = list(subject = as.factor), values_to = "time") %>%
  relocate(subject, id, mode, hotspot, time) %>%
  arrange(subject, id) %T>% print
```
***
# Wilcoxon
***

```{r}
data.long.wil <- data.long %>% arrange(subject, hotspot, mode) %T>% print
```

```{r}
wsrt1 <- wilcox.test(formula=time ~ mode, data=data.long.wil, paired=TRUE, alternative="two.sided")
wsrt1 %>% tidy
```

```{r}
G <- data.long.wil %>% filter(mode == "G") %>% select(time) %T>% print
NG <- data.long.wil %>% filter(mode == "NG") %>% select(time) %T>% print

wsrt2 <- wilcox.test(NG$time, G$time, paired=TRUE, alternative="two.sided")
wsrt2 %>% tidy
```

```{r}
wilcox_test(time ~ mode, data=data.long.wil, paired=TRUE, alternative="less")
```

```{r}
wrst.z <- wilcoxonZ(NG$time, G$time, paired = TRUE, digits = 3)
wrst.z
```

```{r}
wrst.z / sqrt(nrow(NG))
data.long.wil %>% rstatix::wilcox_effsize(formula=time ~ mode, paired=TRUE)
```
```{r}
sign_test(time ~ mode, data=data.long.wil)
```

# Success:

```{r}
data.long.logit <- data.long %>% mutate(success = ifelse(time >= 120, 0, 1)) %T>% print
```

```{r}
tbl1 <- table(data.long.logit$mode, data.long.logit$success)
tbl1
```

```{r}
tbl2 <- prop.table(tbl1) %>% round(2)
tbl2
```

```{r}
chisq_test(tbl1)
```
***
# Graphics
***

```{r}
hist1 <- data.long %>% gghistogram(x = "time", add = "median", rug = TRUE, color = "mode", fill = "mode", 
                          palette = c("#00AFBB", "#E7B800"), bins = 20, xlab="Time to find hotspot", ylab="Count")

hist2 <- data.long %>% gghistogram(x = "time", add = "median", rug = TRUE, color = "mode", fill = "mode", 
                          palette = c("#00AFBB", "#E7B800"), bins = 20, xlab="Time to find hotspot", ylab="Count", facet.by="mode")

ggsave("../res/img/hist_diff.png", hist1, device = "png", dpi=320)
ggsave("../res/img/hist_diff.png", hist2, device = "png", dpi=320)
```

```{r warnings=F}
data.pairwise <- data.long %>% group_by(subject, hotspot) %>%
  summarize(diff = abs(diff(time))) %>%
  ungroup() %T>% print
```

```{r}
data.pairwise %>% gghistogram(x = "diff", add = "median", rug = TRUE, bins = 20, 
                                           color="#FC4E07", fill="#FC4E07", xlab="Paired absolute time-differences", ylab="Count")

ggsave("../res/img/hist_diff.png", device = "png", dpi=320)
```

```{r}
data.long %>% arrange(subject, hotspot) %>% ggdotchart(
  x = "hotspot",
  y = "time",
  color = "mode",
  palette = c("#00AFBB", "#E7B800"),
  sorting = "ascending",
  add = "segments",
  add.params = list(color = "lightgray", size = 1.5),
  group = "mode",
  dot.size = 7,
  label = "time",
  font.label = list(
    color = "white",
    size = 8,
    vjust = 0.5
  ),
  facet.by = "subject",
  xlab="Hotspot",
  ylab="Time to find hotspot",
  ylim = c(0, 130)
)

ggsave("../res/img/full.png", device = "png", dpi=320)
```

```{r}
dat <- data.long %>% Rmisc::summarySE(measurevar="time", groupvars=c("mode", "subject")) %>% print
pd <- position_dodge(0.3)

conflict_prefer("arrange", "plyr")

fig <- dat %>%
  ggplot(aes(x = mode, y = time, color = subject)) + #group=1
  geom_line(position = pd) +
  geom_errorbar(aes(ymin = time - se, ymax = time + se), size = 0.9, width = 0.15, position = pd) +
  geom_point(aes(fill=subject), shape=21, size=3.2, position = pd) +
  theme_classic() +
  labs(x="Guidance mode", y="Time to find hotspot")

fig
ggsave("../res/img/lines.png", device = "png", dpi=320)
#ggplotly(fig)
```

### 2020-06-10 R/Index.Rmd
Update analysis & figs

@@
 if (!require(pacman)) install.packages("pacman")
-pkg_list <- c("tidyverse", "glue", "broom", "magrittr", "prettydoc", "DT", "reshape2", "plotly", "moments", "emmeans", "gtsummary", "lmtest", "table1", "rstatix", "lme4", "ggeffects", "conflicted", "forecast", "Rmisc", "gridExtra", "kableExtra", "magick", "webshot")
+pkg_list <- c("tidyverse", "glue", "broom", "magrittr", "prettydoc", "DT", "reshape2", "plotly", "moments", "emmeans", "gtsummary", "lmtest", "table1", "rstatix", "lme4", "ggeffects", "conflicted", "forecast", "Rmisc", "gridExtra", "kableExtra", "rcompanion")
 

### 2020-06-10 manuscript.tex
Updates from Overleaf

@@
 
+%%TODO: figure description
+%%TODO: split the 4 figures ?
 \begin{figure}
+\includegraphics[width=5.5cm, height=4.1cm,]{res/img/histNG.png}
+\includegraphics[width=6.5cm, keepaspectratio,]{res/img/lines.png}
 \includegraphics[width=5.5cm, keepaspectratio,]{res/img/fig2a.jpg}
 \includegraphics[width=6.5cm, keepaspectratio,]{res/img/fig2b.jpg}
-\caption{(left) Plot of time (x axis) to find target hotspot location for G (‘o’ symbol) vs.
-NG  (‘x’ symbol) conditions, with time-out declared after 120 sec.
-of unsuccessful search.
-Results are shown for each participant (indicated on the y axis and also using color) and at all 10 target locations.
-(right) Histogram plot showing the distribution of NG time minus G time.
-The plot aggregates time differences for all participants over all locations; positive times mean the search was faster with G than with NG.}
+\caption{
+(Upper-left) Histogram of the times to find the hotspots per guidance condition. 
+(Upper-right) Line plot showing the average time to find the hotspot ($\pm{ SE}$) per subject and per guidance condition. 
+(Bottom-left) Time to find target hotspot location for G (‘o’ symbol) vs.
+NG  (‘x’ symbol) conditions.
+The times for each of the 10 hotspots are shown, for each participant (indicated on the y axis and also using color).
+(Bottom-right) Histogram of the differences (paired by hotspot) in time to find the hotspot.}
 \label{fig:res}
@@
 
-We analyzed the differences between times under G vs.~NG using a
-Wilcoxon signed rank test. The resulting p-value is p = 3.8 x 10-8. This
-is a highly statistically significant p-value supporting the rejection
-of the null hypothesis that the distribution of times under G is the
-same as the distribution of times under NG.
+We used a Wilcoxon Signed Rank test to compare the differences between G and NG times, paired by hotspot within each participant.
+The result indicated that participants performed significantly better in the guided condition (med = 14,5) than in the non guided one (med = 120), with T=815, z=5.45, $p < 0.001$, and an effect size of V = 0.86.
 
-The following System Usability Scale (SUS) scores were calculated:
-P1=65, P2=75, P3=77.5, for an average of 72.5. An SUS score of 68 is
-considered ``average'' while scores in the mid- to high 70s are
-considered ``good'' (\url{https://measuringu.com/interpret-sus-score/}).
-Overall, the SUS scores suggest that the guidance system is usable but
-needs improvement.
+The following System Usability Scale (SUS) scores were calculated: P1=65, P2=75, P3=77.5, for an average of 72.5. 
+An SUS score of 68 is considered ``average'' while scores in the mid- to high 70s are considered ``good'' (\url{https://measuringu.com/interpret-sus-score/}).
+Overall, the SUS scores suggest that the guidance system is usable but needs improvement.
 
-Next we summarize the qualitative feedback we obtained in the
-semi-structured interviews. All participants who provided qualitative
-feedback (P1, P2 and P3) acknowledged the benefit of the audio guidance
-that was provided, which they found intuitive and clearly superior to
-finding hotspots without guidance. However, P1 and P2 felt that the
-limited spatial accuracy of the CamIO system was the biggest problem
-that needed to be addressed; for instance, in some cases the system
-announced the target had been reached even though the stylus tip was
-hovering about 1 cm above the truck surface. (This spatial inaccuracy
+Next we summarize the qualitative feedback we obtained in the semi-structured interviews. 
+All participants who provided qualitative feedback (P1, P2 and P3) acknowledged the benefit of the audio guidance that was provided, which they found intuitive and clearly superior to finding hotspots without guidance.
+However, P1 and P2 felt that the limited spatial accuracy of the CamIO system was the biggest problem that needed to be addressed; for instance, in some cases the system announced the target had been reached even though the stylus tip was hovering about 1 cm above the truck surface.
+(This spatial inaccuracy
 results from noise in the board pose and stylus tip location estimates.)
-P3 expressed frustration with time lag, i.e., the voice feedback tended
-to lag behind the current location of the stylus, even when she was
-moving the stylus slowly; she also felt that the speed of the repetitive
-speech feedback made her feel rushed.
+P3 expressed frustration with time lag, i.e., the voice feedback tended to lag behind the current location of the stylus, even when she was moving the stylus slowly; she also felt that the speed of the repetitive speech feedback made her feel rushed.
 
 When asked what improvements the participants wanted, P1 suggested
-equipping the stylus with a button that turns on the guidance feedback,
-and the addition of multiple vibrators on the stylus that could provide
-haptic feedback indicating which way the stylus should be moved. P2
-suggested adding audio information similar to the earlier guidance
-system that the authors experimented with (Sec. 3), which would use a
-combination of tempo, pitch and volume changes to signal when the stylus
-gets closer to the target location. P3 expressed the desire that CamIO
-work with any ordinary pen or pencil, rather than requiring the special
-stylus we used.
+equipping the stylus with a button that turns on the guidance feedback, and the addition of multiple vibrators on the stylus that could provide haptic feedback indicating which way the stylus should be moved.
+P2 suggested adding audio information similar to the earlier guidance system that the authors experimented with (Sec.
+3), which would use a combination of tempo, pitch and volume changes to signal when the stylus gets closer to the target location.
+P3 expressed the desire that CamIO work with any ordinary pen or pencil, rather than requiring the special stylus we used.
 

### 2020-06-10 R/ICCHP20.Rmd
Update graphics

@@
 
+conflict_prefer("arrange", "plyr")
 conflict_prefer("mutate", "dplyr")
+conflict_prefer("rename", "dplyr")
 conflict_prefer("select", "dplyr")
@@
 options(knitr.table.format = ifelse(knitr::is_latex_output(), "latex", "pandoc"), knitr.kable.NA = "")
+
+hs.order <- c("1","2","3","4","5","6","7","8","9","10")
 ```
@@
   mutate(mode = ifelse(mode == "N", "NG", "G")) %>%
-  mutate(across(c(hotspot, mode), as.factor))
+  rename("Condition" = mode, "Hotspot" = hotspot) %>%
+  mutate(across(c(Hotspot, Condition), as.factor)) %T>% print
 
 data.long <- data.wide %>%
-  mutate(id = seq_along(hotspot), .before = hotspot) %>%
-  pivot_longer(cols = "P1":"P4", names_to="subject", names_transform = list(subject = as.factor), values_to = "time") %>%
-  relocate(subject, id, mode, hotspot, time) %>%
-  arrange(subject, id) %T>% print
+  mutate(id = seq_along(Hotspot), .before = Hotspot) %>%
+  pivot_longer(cols = "P1":"P4", names_to="Participant", names_transform = list(Participant = as.factor), values_to = "Time") %>%
+  relocate(Participant, id, Condition, Hotspot, Time) %>%
+  #mutate(Hotspot = factor(Hotspot, levels = hs.order, ordered = TRUE)) %>%
+  dplyr::arrange(Participant, id) %T>% print
 ```
@@
 ```{r}
-data.long.wil <- data.long %>% arrange(subject, hotspot, mode) %T>% print
+data.long.wil <- data.long %>% dplyr::arrange(Participant, Hotspot, Condition) %T>% print
 ```
@@
 ```{r}
-wsrt1 <- wilcox.test(formula=time ~ mode, data=data.long.wil, paired=TRUE, alternative="two.sided")
+wsrt1 <- wilcox.test(formula=Time ~ Condition, data=data.long.wil, paired=TRUE, alternative="two.sided")
 wsrt1 %>% tidy
@@
 ```{r}
-G <- data.long.wil %>% filter(mode == "G") %>% select(time) %T>% print
-NG <- data.long.wil %>% filter(mode == "NG") %>% select(time) %T>% print
+G <- data.long.wil %>% filter(Condition == "G") %>% select(Time) %T>% print
+NG <- data.long.wil %>% filter(Condition == "NG") %>% select(Time) %T>% print
 
-wsrt2 <- wilcox.test(NG$time, G$time, paired=TRUE, alternative="two.sided")
+wsrt2 <- wilcox.test(NG$Time, G$Time, paired=TRUE, alternative="two.sided")
 wsrt2 %>% tidy
@@
 ```{r}
-wilcox_test(time ~ mode, data=data.long.wil, paired=TRUE, alternative="less")
+wilcox_test(Time ~ Condition, data=data.long.wil, paired=TRUE, alternative="less")
 ```
@@
 ```{r}
-wrst.z <- wilcoxonZ(NG$time, G$time, paired = TRUE, digits = 3)
+wrst.z <- wilcoxonZ(NG$Time, G$Time, paired = TRUE, digits = 3)
 wrst.z
@@
 wrst.z / sqrt(nrow(NG))
-data.long.wil %>% rstatix::wilcox_effsize(formula=time ~ mode, paired=TRUE)
+data.long.wil %>% rstatix::wilcox_effsize(formula=Time ~ Condition, paired=TRUE)
 ```
+
 ```{r}
-sign_test(time ~ mode, data=data.long.wil)
+sign_test(Time ~ Condition, data=data.long.wil)
 ```
@@
 ```{r}
-data.long.logit <- data.long %>% mutate(success = ifelse(time >= 120, 0, 1)) %T>% print
+data.long.logit <- data.long %>% mutate(Success = ifelse(Time >= 120, 0, 1)) %T>% print
 ```
@@
 ```{r}
-tbl1 <- table(data.long.logit$mode, data.long.logit$success)
+tbl1 <- table(data.long.logit$Condition, data.long.logit$Success)
 tbl1
@@
 ```{r}
-hist1 <- data.long %>% gghistogram(x = "time", add = "median", rug = TRUE, color = "mode", fill = "mode", 
-                          palette = c("#00AFBB", "#E7B800"), bins = 20, xlab="Time to find hotspot", ylab="Count")
+hist1 <-
+  data.long %>% gghistogram(
+    x = "Time",
+    rug = T,
+    add = "median",
+    color = "Condition",
+    fill = "Condition",
+    alpha=0.75,
+    palette = c("#00AFBB", "#f48806"),
+    bins = 20,
+    xticks.by=20,
+    xlab = "Time to find hotspot",
+    ylab = "Count"
+  )
 
-hist2 <- data.long %>% gghistogram(x = "time", add = "median", rug = TRUE, color = "mode", fill = "mode", 
-                          palette = c("#00AFBB", "#E7B800"), bins = 20, xlab="Time to find hotspot", ylab="Count", facet.by="mode")
+hist2 <-
+  data.long %>% gghistogram(
+    x = "Time",
+    rug = T,
+    add = "median",
+    color = "Condition",
+    fill = "Condition",
+    alpha=0.75,
+    palette = c("#00AFBB", "#f48806"),
+    bins = 20,
+    xticks.by=20,
+    xlab = "Time to find hotspot",
+    ylab = "Count",
+    facet.by = "Condition"
+  )
 
-ggsave("../res/img/hist_diff.png", hist1, device = "png", dpi=320)
-ggsave("../res/img/hist_diff.png", hist2, device = "png", dpi=320)
+hist1
+ggsave("../res/img/hist1.png", hist1, device = "png", dpi=320)
+hist2
+ggsave("../res/img/hist2.png", hist2, device = "png", dpi=320)
 ```
@@
 ```{r warnings=F}
-data.pairwise <- data.long %>% group_by(subject, hotspot) %>%
-  summarize(diff = abs(diff(time))) %>%
+data.pairwise <- data.long %>% 
+  dplyr::arrange(Participant, Hotspot, desc(Condition)) %>%
+  group_by(Participant, Hotspot) %>%
+  summarize(Diff = -1*diff(Time)) %>%
   ungroup() %T>% print
@@
 ```{r}
-data.pairwise %>% gghistogram(x = "diff", add = "median", rug = TRUE, bins = 20, 
-                                           color="#FC4E07", fill="#FC4E07", xlab="Paired absolute time-differences", ylab="Count")
+data.pairwise %>% gghistogram(
+  x = "Diff",
+  bins = 20,
+  rug = T,
+  color = "#0c8c1a",
+  fill = "#0c8c1a",
+  xlab = "Paired time-differences",
+  ylab = "Count",
+  xticks.by=20
+)
 
-ggsave("../res/img/hist_diff.png", device = "png", dpi=320)
+ggsave("../res/img/hist_diff.png",
+       device = "png",
+       dpi = 320)
 ```
 
-```{r}
-data.long %>% arrange(subject, hotspot) %>% ggdotchart(
-  x = "hotspot",
-  y = "time",
-  color = "mode",
-  palette = c("#00AFBB", "#E7B800"),
-  sorting = "ascending",
+```{r warnings=T}
+test <- data.long %>% 
+  select(Hotspot, Participant, Condition, Time) %>%
+  group_by(Condition) %>% 
+  mutate(Hotspot = factor(Hotspot, levels=hs.order)) %>% 
+  ungroup %>%
+  arrange(Participant, Hotspot) %T>% print
+
+#data.long$Hotspot <- factor(data.long$Hotspot, levels = hs.order)
+
+fig.full <- ggdotchart(
+  test,
+  x = "Hotspot",
+  y = "Time",
+  color = "Condition",
+  #shape = "Condition",
+  palette = c("#00AFBB", "#f48806"),
   add = "segments",
   add.params = list(color = "lightgray", size = 1.5),
-  group = "mode",
-  dot.size = 7,
-  label = "time",
+  group = "Condition",
+  dot.size = 4,
+  #label = "Time",
   font.label = list(
@@
   ),
-  facet.by = "subject",
-  xlab="Hotspot",
-  ylab="Time to find hotspot",
-  ylim = c(0, 130)
+  facet.by = "Participant",
+  xlab = "Hotspot",
+  ylab = "Time to find hotspot",
+  ylim = c(0, 130),
+  yticks.by = 20
 )
 
-ggsave("../res/img/full.png", device = "png", dpi=320)
+fig.full
+ggsave("../res/img/full.png", fig.full, device = "png", dpi=320)
 ```
@@
 ```{r}
-dat <- data.long %>% Rmisc::summarySE(measurevar="time", groupvars=c("mode", "subject")) %>% print
+dat <- data.long %>% Rmisc::summarySE(measurevar="Time", groupvars=c("Condition", "Participant")) %>% print
 pd <- position_dodge(0.3)
 
-conflict_prefer("arrange", "plyr")
-
 fig <- dat %>%
-  ggplot(aes(x = mode, y = time, color = subject)) + #group=1
+  ggplot(aes(x = Condition, y = Time, color = Participant)) + #group=1
   geom_line(position = pd) +
-  geom_errorbar(aes(ymin = time - se, ymax = time + se), size = 0.9, width = 0.15, position = pd) +
-  geom_point(aes(fill=subject), shape=21, size=3.2, position = pd) +
+  geom_errorbar(aes(ymin = Time - se, ymax = Time + se), size = 0.9, width = 0.15, position = pd) +
+  geom_point(aes(fill=Participant), shape=21, size=3.2, position = pd) +
   theme_classic() +
-  labs(x="Guidance mode", y="Time to find hotspot")
+  theme(axis.text.x = element_text(size=12, color="black"),
+        axis.text.y = element_text(size=12, color="black")) +
+  labs(x="Guidance condition", y="Time to find hotspot", ylim=c(0,120))
 

### 2020-06-10 manuscript.tex

@@
 
-%%TODO: figure description
-%%TODO: split the 4 figures ?
 \begin{figure}
-\includegraphics[width=5.5cm, height=4.1cm,]{res/img/histNG.png}
-\includegraphics[width=6.5cm, keepaspectratio,]{res/img/lines.png}
-\includegraphics[width=5.5cm, keepaspectratio,]{res/img/fig2a.jpg}
-\includegraphics[width=6.5cm, keepaspectratio,]{res/img/fig2b.jpg}
+\includegraphics[width=6cm, keepaspectratio,]{res/img/hist1.png}
+\includegraphics[width=6cm, keepaspectratio,]{res/img/lines.png}
+\includegraphics[width=6cm, keepaspectratio,]{res/img/hist_diff.png}
+\includegraphics[width=6cm, keepaspectratio,]{res/img/full.png}
 \caption{
-(Upper-left) Histogram of the times to find the hotspots per guidance condition. 
-(Upper-right) Line plot showing the average time to find the hotspot ($\pm{ SE}$) per subject and per guidance condition. 
-(Bottom-left) Time to find target hotspot location for G (‘o’ symbol) vs.
-NG  (‘x’ symbol) conditions.
-The times for each of the 10 hotspots are shown, for each participant (indicated on the y axis and also using color).
-(Bottom-right) Histogram of the differences (paired by hotspot) in time to find the hotspot.}
+\emph{(Upper-left)} Histogram of the times to find the hotspots per guidance condition. 
+\emph{(Upper-right)} Line plot showing the average time to find the hotspot ($\pm{ SE}$) per subject and per guidance condition. 
+\emph{(Bottom-left)} Histogram of the differences (paired by hotspot) in time to find the hotspot.
+\emph{(Bottom-right)} Time to find target hotspot location for each participant, colored by guidance mode.}
 \label{fig:res}

### 2020-06-10 R/ICCHP20.Rmd
Finally ...

@@
   relocate(Participant, id, Condition, Hotspot, Time) %>%
-  #mutate(Hotspot = factor(Hotspot, levels = hs.order, ordered = TRUE)) %>%
+  #mutate(Hotspot = factor(Hotspot, levels = hs.order)) %>%
   dplyr::arrange(Participant, id) %T>% print
@@
 ```{r}
-hist1 <-
 data.long %>% gghistogram(
@@
   fill = "Condition",
-    alpha=0.75,
+  alpha = 0.75,
   palette = c("#00AFBB", "#f48806"),
   bins = 20,
-    xticks.by=20,
-    xlab = "Time to find hotspot",
-    ylab = "Count"
-  )
-
-hist2 <-
-  data.long %>% gghistogram(
-    x = "Time",
-    rug = T,
-    add = "median",
-    color = "Condition",
-    fill = "Condition",
-    alpha=0.75,
-    palette = c("#00AFBB", "#f48806"),
-    bins = 20,
-    xticks.by=20,
+  xticks.by = 20,
   xlab = "Time to find hotspot",
@@
 
-hist1
-ggsave("../res/img/hist1.png", hist1, device = "png", dpi=320)
-hist2
-ggsave("../res/img/hist2.png", hist2, device = "png", dpi=320)
+ggsave("../res/img/hist.png", device = "png", dpi = 320)
 ```
@@
 ```{r warnings=T}
-test <- data.long %>% 
-  select(Hotspot, Participant, Condition, Time) %>%
-  group_by(Condition) %>% 
-  mutate(Hotspot = factor(Hotspot, levels=hs.order)) %>% 
-  ungroup %>%
-  arrange(Participant, Hotspot) %T>% print
-
-#data.long$Hotspot <- factor(data.long$Hotspot, levels = hs.order)
-
 fig.full <- ggdotchart(
-  test,
+  data.long,
   x = "Hotspot",
@@
   add.params = list(color = "lightgray", size = 1.5),
-  group = "Condition",
+  group = "Hotspot",
   dot.size = 4,
@@
   yticks.by = 20
-)
+) #+ scale_x_discrete(labels=hs.order)
 

### 2020-06-17 R/ICCHP20.Rmd
Update Python notebook

@@
 ```{r}
-wsrt1 <- wilcox.test(formula=Time ~ Condition, data=data.long.wil, paired=TRUE, alternative="two.sided")
+wsrt1 <- wilcox.test(formula=Time ~ Condition, data=data.long.wil, paired=TRUE, alternative="two.sided", exact=FALSE, correct=FALSE)
 wsrt1 %>% tidy
+wsrt1$statistic
 ```
@@
 
-wsrt2 <- wilcox.test(NG$Time, G$Time, paired=TRUE, alternative="two.sided")
+wsrt2 <- wilcox.test(NG$Time, G$Time, paired=TRUE, alternative="two.sided", exact=FALSE, correct=FALSE)
 wsrt2 %>% tidy
@@
 ```{r}
-wrst.z / sqrt(nrow(NG))
-data.long.wil %>% rstatix::wilcox_effsize(formula=Time ~ Condition, paired=TRUE)
+#wrst.z / sqrt(nrow(NG))
+data.long.wil %>% rstatix::wilcox_effsize(formula=Time ~ Condition, paired=TRUE, alternative="less")
 ```

### 2020-06-17 manuscript.tex
Updates from Overleaf

--- 2020-06-09 manuscript.tex
@@
 \usepackage{hyperref}
+\usepackage{draftwatermark}
+
+\SetWatermarkText{Preprint – to be published in ICCHP 2020}
+\SetWatermarkColor[gray]{0.5}
+\SetWatermarkFontSize{1cm}
+\SetWatermarkAngle{90}
+\SetWatermarkHorCenter{2cm}
 
@@
 
-\title{An Audio-Based 3D Spatial Guidance AR System for Blind Users\thanks{JMC, BB and HS were supported by NIH grant 5R01EY025332 and NIDILRR grant 90RE5024-01-00.
-We thank Dr.
-Roberto Manduchi, Dr.
-Natela Shanidze, Dr.
-Santani Teng and Dr.
-Ali Cheraghi for helpful suggestions about the experiments.}}
+\title{An Audio-Based 3D Spatial Guidance AR System for Blind Users}
 
@@
 by the user to point at locations of interest on an object (called
-hotspots). When the user points at a hotspot with the stylus tip,
+{\em hotspots}). When the user points at a hotspot with the stylus tip,
 relevant audio information, such as text-to-speech or other sounds, is
@@
 
-Experiments with blind participants have demonstrated the feasibility of
-using CamIO to access existing audio labels. Moreover, our approach
-allows a visually impaired user to create audio labels independently
-\cite{coughlanTowardsAccessibleAudioLabeling2020}. However, an ability
-that CamIO still lacks is the ability to provide real-time guidance to a
-hotspot. Without such guidance, the user must search the entire surface
-of an object with the stylus tip until they hear the hotspot they are
-seeking. This need is particularly acute in cases where the surface is
-lacking in tactile cues, such as the touch screen buttons on a microwave
-oven \cite{coughlanTowardsAccessibleAudioLabeling2020}.
+Experiments with blind participants have demonstrated the feasibility of using CamIO to access existing audio labels. 
+Moreover, our approach allows a visually impaired user to create audio labels independently \cite{coughlanTowardsAccessibleAudioLabeling2020}.
+However, an ability that CamIO still lacks is the ability to provide real-time guidance to a hotspot.
+Without such guidance, the user must search the entire surface of an object with the stylus tip until they hear the hotspot they are seeking. 
+This need is particularly acute in cases where the surface is lacking in tactile cues, such as the touch screen buttons on a microwave oven \cite{coughlanTowardsAccessibleAudioLabeling2020}.
 
-Guidance not only saves the user time in exploring large or intricate
-objects, or objects lacking in tactile features, but can enable entirely
-new ways of interacting with objects. For instance, a geography learning
-app could ask a student to point to a specific city on a world globe,
-and if the student has trouble finding it, the app could guide them to
-it in an intuitive way. Other examples could include guiding a user to a
-specific button on an appliance (such as a microwave) to perform a
-desired function, guiding a student to a specific element on the
-periodic table, or even guiding a repairperson to find a part of a
-machine with limited visibility. Our guidance system is inspired by
-recent work on auditory and haptic displays for 3D spatial guidance
-\cite{mayAuditoryDisplaysFacilitate2019,guezou-philippePrototypingEvaluatingSensory2018},
-which provide real-time audio/haptic feedback to guide a user towards a
-target in a virtual environment. We initially developed an interface
-similar to \cite{mayAuditoryDisplaysFacilitate2019} for CamIO, but after
-preliminary testing we devised a simpler and more effective interface
-for our application, which we evaluated experimentally with four blind
-participants.
+Guidance not only saves the user time in exploring large or intricate objects, or objects lacking in tactile features, but can enable entirely new ways of interacting with objects. 
+For instance, a geography learning app could ask a student to point to a specific city on a world globe, and if the student has trouble finding it, the app could guide them to it in an intuitive way. 
+Other examples could include guiding a user to a specific button on an appliance (such as a microwave oven) to perform a desired function, guiding a student to a specific element on the periodic table, or even guiding a repair-person to find a part of a machine with limited visibility.
+Our guidance system is inspired by recent work on auditory and haptic displays for 3D spatial guidance \cite{mayAuditoryDisplaysFacilitate2019,guezou-philippePrototypingEvaluatingSensory2018,miesenberger_augmented_2018}, which provide real-time audio/haptic feedback to guide a user towards a target in a virtual environment.
+We initially developed an interface similar to \cite{mayAuditoryDisplaysFacilitate2019} for CamIO, but after preliminary testing we devised a simpler and more effective interface for our application, which we evaluated experimentally with four blind participants.
 
@@
 to provide audio labels for rigid 3D objects. The rigid object of
-interest is mounted on a board, which is a printed barcode pattern
-(Fig. \ref{fig:exp}) that allows CamIO to estimate the object's pose (3D
+interest is mounted on a {\em board}, which is a flat printed barcode pattern
+(Fig.~\ref{fig:exp}) that allows CamIO to estimate the object's pose (3D
 translation and 3D orientation) relative to the camera. The iPhone is
 typically mounted on a tripod or a gooseneck cellphone holder
-Fig. ref{fig:exp} to obtain a clear view of the board and object, but in
+(Fig.~\ref{fig:exp}(left)) to obtain a clear view of the board and object, but in
 some cases \cite{coughlanTowardsAccessibleAudioLabeling2020} users may
 hold the iPhone by hand. The user points the tip of a passive wooden
-stylus Fig. \ref{fig:exp}, covered with another barcode pattern, to different
+{\em stylus} (Fig.~\ref{fig:exp}), covered with another barcode pattern, to different
 locations on an object. CamIO estimates the stylus tip's 3D location
 relative to the camera and uses the pose estimate of the board to
-determine the stylus tip's 3D location relative to the object. Whenever
+determine the stylus tip's 3D location relative to the object. (In other words, hotspot locations are
+defined relative to the board, which is mounted rigidly to the object.) Whenever
 the tip is close enough to a pre-defined hotspot, it triggers audio
-information about the hotspot. (In other words, hotspot locations are
-defined relative to the board, which is mounted rigidly to the object.)
+information about the hotspot.
 
 A second stylus (not shown, identical in form but distinct from the one
-shown in Fig. \ref{fig:exp}) is used to create new hotspots, and allows the
+shown in Fig.~\ref{fig:exp}) is used to create new hotspots, and allows the
 user to make an audio recording to associate with each new hotspot. Our
@@
 
-We build on the auditory display described in
-\cite{mayAuditoryDisplaysFacilitate2019}, which is intended for
-audio-based 3D targeting of locations in a large virtual workspace (such
-as a kitchen). Accordingly, the first incarnation of our guidance
-approach used a repeating monaural sound, whose tempo indicated how far
-the stylus tip is from the target, and with a pitch (low, medium or
-high) that indicated whether the tip needed to move up or down (or else
-was aligned properly in the vertical direction).
+We build on the auditory display described in \cite{mayAuditoryDisplaysFacilitate2019}, which is intended for
+audio-based 3D targeting of locations in a large virtual workspace (such as a kitchen). 
+Accordingly, the first incarnation of our guidance approach used a repeating monaural sound, whose tempo indicated how far
+the stylus tip is from the target, and with a pitch (low, medium or high) that indicated whether the tip needed to move up or down (or else was aligned properly in the vertical direction).
 
-Informal tests with two blind participants (one female of age 42, who is
-also participant P1 in the experiment described in the next section, and
-one male of age 73) showed that this guidance worked satisfactorily in
-some cases. However, we found that the feedback was too indirect: the
-participants didn't always know how to move the stylus to increase the
-tempo. As a result, the resulting guidance process was unacceptably
-slow. Thus we devised a second, more direct guidance approach: verbal
-directions telling the user to move the stylus along the cardinal
-directions (up, down, left, right, forward or back). While the optimal
-direction to move the stylus tip towards the hotspot might entail a
-combination of these directions (e.g., simultaneously move left and up),
-for simplicity we issue directions along only one cardinal direction at
-a time -- the direction that will most improve the alignment to the
-target.
+Informal tests with two blind participants (one female of age 42, who is also participant P1 in the experiment described in the next section, and one male of age 73) showed that this guidance worked satisfactorily in some cases.
+However, we found that the feedback was too indirect: the participants didn't always know {\em how} to move the stylus to increase the tempo. 
+As a result, the resulting guidance process was unacceptably slow. 
+Thus we devised a second, more direct guidance approach: verbal directions telling the user to move the stylus along the cardinal directions (up, down, left, right, forward or back). 
+While the optimal direction to move the stylus tip towards the hotspot might entail a combination of these directions (e.g., simultaneously move left and up), for simplicity we issue directions along only {\em one} cardinal direction at a time -- the direction that will most improve the alignment to the target.
 
-We define the guidance interface as follows. The board (barcode pattern
-beneath the truck, see Fig. \ref{fig:exp}) defines an xyz coordinate frame,
-with \(+x\) = right, \(-x\) = left, \(+y\) = forward (away from the
-user), \(-y\) = back (towards the user), \(+z\) = up and \(-z\) = down.
-Let the desired target hotspot location be denoted \((x*, y*, z*)\), and
-the current location of the stylus tip be \((x,y,z)\). Define the error
-vector \(e = (x-x*, y-y*, z-z*)\), and let index \(a = argmax_i |e_i|\),
-i.e., a is the index (1, 2 or 3, corresponding to x, y or z,
-respectively) of the entry in e with the highest absolute value. In
-other words, a indicates which coordinate (x, y or z) is most discrepant
-from the target coordinate location \((x*, y*, z*)\). Then the
-appropriate direction is issued, e.g., ``left'' if \(a = 1\) and
-\(x > x*\), ``right'' if \(a = 1\) and \(x < x*\), ``up'' if \(a = 3\)
-and \(z < z*\), etc. Directions are announced roughly twice per second
-whenever the stylus is visible. The target hotspot is announced when the
-stylus gets to within approximately 1 cm from the hotspot. No audio
-feedback is issued when the stylus is not visible.
+We define the guidance interface as follows.
+The board (barcode pattern beneath the truck, see Fig.~\ref{fig:exp}) defines an xyz coordinate frame, with \(+x\) = right, \(-x\) = left, \(+y\) = forward (away from the user), \(-y\) = back (towards the user), \(+z\) = up and \(-z\) = down.
+Let the desired target hotspot location be denoted \((x^*, y^*, z^*)\), and the current location of the stylus tip be \((x,y,z)\). 
+Define the error vector \(e = (x-x^*, y-y^*, z-z^*)\), and let index \(a = argmax_i |e_i|\), i.e., $a$ is the index (1, 2 or 3, corresponding to x, y or z, respectively) of the entry in $e$ with the highest absolute value.
+In other words, $a$ indicates which coordinate (x, y or z) is most discrepant from the target coordinate location \((x^*, y^*, z^*)\). 
+Then the appropriate direction is issued, e.g., ``left'' if \(a = 1\) and \(x > x^*\), ``right'' if \(a = 1\) and \(x < x^*\), ``up'' if \(a = 3\) and \(z < z^*\), etc. 
+Directions are announced roughly twice per second whenever the stylus is visible. The target hotspot is announced when the stylus gets to within approximately 1 cm from the hotspot. 
+No audio feedback is issued when the stylus is not visible.
 
 \hypertarget{exp}{%
-\section{Experiment with Blind Participants}\label{exp}}
+\section{Methods}\label{exp}}
 
-We conducted an experiment to assess the feasibility of the spatial
-guidance feedback and to determine whether the feedback speeds up the
-target search. The experiment Fig. \ref{fig:exp} compared how much time a blind participant needed to find hotspots (points of interest) on a 3D
-object with spatial guidance vs.~without it. The object was a large toy
-fire truck (27 cm x 12 cm x 14 cm), with K=10 distinct hotspot locations
-defined in advance. The hotspots were tactilely salient features such as
-lights, hooks or other features that protruded from the truck. Some
-hotspots were in recessed locations on the truck, such as the shift
-stick and steering wheel (Fig. \ref{fig:exp}) in the front cabin, or on the
-side of the truck facing away from the participant; these hotspots were
-difficult to find and access, whether by fingertip or by stylus. In
-fact, the difficulty of accessing such locations highlights an advantage
-of the CamIO stylus, whose tip need not be visible to the camera. This
-allows the user to explore recessed areas or areas that are otherwise
-invisible to the camera.
+We conducted an experiment to assess the feasibility of the spatial guidance feedback and to determine whether the feedback speeds up the target search. 
+The experiment (see setup in Fig.~\ref{fig:exp} (left)) compared how much time a blind participant needed to find hotspots (points of interest) on a 3D object with spatial guidance vs.~without it. 
+The object was a large toy fire truck (27 cm x 12 cm x 14 cm), with K=10 distinct hotspot locations defined in advance. 
+The hotspots were tactilely salient features such as lights, hooks or other features that protruded from the truck. 
+Some hotspots were in recessed locations on the truck, such as the shift stick and steering wheel (Fig.~\ref{fig:exp} (right)) in the front cabin, or on the side of the truck facing away from the participant; these hotspots were difficult to find and access, whether by fingertip or by stylus.
+In fact, the difficulty of accessing such locations highlights an advantage of the CamIO stylus, whose tip need not be visible to the camera. 
+This allows the user to explore recessed areas or areas that are otherwise invisible to the camera.
 
 \begin{figure}
-\includegraphics[width=0.55\linewidth]{res/img/fig1a}
-\includegraphics[width=0.45\linewidth]{res/img/fig1b} 
+\includegraphics[width=0.58\linewidth,height=4.5cm]{res/img/fig1a}
+\includegraphics[width=0.42\linewidth,height=4.5cm]{res/img/fig1b} 
+%\includegraphics[height=4.5cm\linewidth]{res/img/fig1a}
+%\includegraphics[height=4.5cm\linewidth]{res/img/fig1b} 
 \caption{(left) Blind participant uses CamIO to explore a toy fire truck.
-The participant holds the CamIO stylus (a barcode pattern printed on a square wooden dowel with a pointy tip) to point to possible hotspots on the truck.
-The truck is glued to the board, a piece of paper with a separate barcode pattern, which is taped to the table.
+The participant holds the CamIO {\em stylus} (a barcode pattern printed on a square wooden dowel with a pointy tip) to point to possible hotspots on the truck.
+The truck is glued to the {\em board} (a piece of paper with a separate barcode pattern), which is taped to the table.
 An iPhone 8 running CamIO is held by a gooseneck cellphone holder to capture a view of the truck, board and stylus.
@@
 
-Each hotspot location was searched for under two feedback conditions: G
-(guidance provided) and NG (no guidance). The participant was given up
-to 120 sec to find each hotspot, after which a time-out was declared.
-In order to minimize the influence of any possible learning effects,
-which could make it easier to find a hotspot the second time it was
-searched for, we randomized the sequence of hotspots and alternated
-between NG and G trials. This resulted in a sequence of 2K = 20 trials,
-starting with an NG trial, the same sequence for each participant.
+Each of the 10 hotspots was searched for under both feedback conditions: G (guidance provided) and NG (no guidance). 
+In order to minimize the influence of any possible learning effects, which could make it easier to find a hotspot the second time it was searched for, we randomized the sequence of hotspots, and alternated between NG and G trials.
+With 10 trials in both guidance conditions (G and NG), this resulted in a within-subject factorial design with an identical sequence of $2K = 20$ trials per participant.
+The participant was given up to 120 sec. to find each hotspot, after which a time-out was declared.
 
-We had a total of N=4 blind participants in our experiment (2 male/2
-female, ages from 26-42). Participants P1, P3 and P4 were already
-familiar with the CamIO system (P4 is a co-author of this paper) while
-P2 was unfamiliar with it. None of the participants was told the study
-design before their trials, and we didn't tell them whether or not any
-locations would be repeated in the 20 trials. After obtaining IRB
-consent, we reviewed how CamIO works if necessary and demonstrated the
-audio guidance feature. The experimenter launched CamIO on an iPhone 8
-and positioned it on a gooseneck cellphone holder so that the camera
-captured the entire truck and board. Each participant spent a short time
-practicing with the guidance feature, with help provided by the
-experimenter as needed. (The experimenter chose practice hotspots that
-were different from the pre-defined K=10 hotspots.) In the practice
-phase, the experimenter explained that barriers sometimes block the user
-from moving in the suggested guidance direction. For instance, if in
-Fig. \ref{fig:exp} the system directs the participant to move ``forward''
-towards the left side mirror, the stylus will soon hit the cabin
-roof/wall as it moves towards the mirror; the participant must find ways
-around such obstacles to find the target hotspot.
+We had a total of N=4 blind participants in our experiment (2 males/2 females, ages from 26-42).
+Participants P1, P3 and P4 were already familiar with the CamIO system (P4 is a co-author of this paper) while P2 was unfamiliar with it. 
+None of the participants was told the study design before their trials, and we didn't tell them whether or not any locations would be repeated in the 20 trials.
 
-Next each participant performed the formal experimental trials, which
-consisted of the 2K=20 trial sequence described above. This sequence
-included 10 trials of both conditions (G and NG), for a total of 20
-trials per participant.
+After obtaining IRB consent, we reviewed how CamIO works if necessary and demonstrated the audio guidance feature. 
+The experimenter launched CamIO on an iPhone 8 and positioned it on a gooseneck cellphone holder so that the camera captured the entire truck and board. 
+Each participant spent a short time practicing with the guidance feature (on different hotspots than the one used for the evaluation), with help provided by the experimenter as needed.
+
+In the practice phase, the experimenter explained that barriers sometimes block the user from moving in the suggested guidance direction. 
+For instance, if in Fig.~\ref{fig:exp}(right) the system directs the participant to move ``forward'' towards the left side mirror, the stylus will soon hit the cabin roof/wall as it moves towards the mirror; the participant must find ways around such obstacles to find the target hotspot.
+
+Next each participant performed the formal experimental trials, which consisted of the 20-trial sequence described above. 
 We  placed a visual barrier between the participant and the object while the experimenter defined the hotspot in each trial to prevent participants from using any residual vision to see the location of the hotspot.
-After completing the 20 search trials, we
-administered a System Usability Scale (SUS) \cite{brooke1996sus} to
-estimate the usability of the guidance system. Finally, we conducted a
-semi-structured interview asking what participants liked and disliked
-about the system, what needed to be improved, and how they might
-envision an ideal guidance system. Since P4 is a co-author on this
-paper, we did not administer the SUS or semi-structured interview to
-him.
+After completing the 20 search trials, we administered a System Usability Scale (SUS) \cite{brooke1996sus} to estimate the usability of the guidance system.
+Finally, we conducted a semi-structured interview asking what participants liked and disliked about the system, what needed to be improved, and how they might envision an ideal guidance system.
+Since P4 is a co-author on this paper, we did not administer the SUS or semi-structured interview to him.
 
 \hypertarget{results}{%
-\section{Experimental Results and Analysis}\label{results}}
+\section{Results and Discussion}\label{results}}
+%%%
 
-We conducted an experiment to assess the feasibility of the spatial
-guidance feedback and to determine whether the feedback speeds up the
-target search. The experiment (Fig. \ref{fig:exp}) compared how much time a
-blind participant needed to find hotspots (points of interest) on a 3D
-object with spatial guidance vs.~without it. The object was a large toy
-fire truck (27 cm x 12 cm x 14 cm), with K=10 distinct hotspot locations
-defined in advance. The hotspots were tactilely salient features such as
-lights, hooks or other features that protruded from the truck. Some
-hotspots were in recessed locations on the truck, such as the shift
-stick and steering wheel (Fig. \ref{fig:exp}) in the front cabin, or on the
-side of the truck facing away from the participant; these hotspots were
-difficult to find and access, whether by fingertip or by stylus. In
-fact, the difficulty of accessing such locations highlights an advantage
-of the CamIO stylus, whose tip need not be visible to the camera. This
-allows the user to explore recessed areas or areas that are otherwise
-invisible to the camera.
+Fig.'s~\ref{fig:res-ul}-\ref{fig:res-full} plot the times needed to find the hotspot in both guidance conditions, and Fig.~\ref{fig:res-full} shows the raw times of each participant for each condition and hotspot.
+All times are in seconds; times reported as 120 secs.
+correspond to participants failing to find the target in the allotted time.
 
 \begin{figure}
-\includegraphics[width=5.5cm, keepaspectratio,]{res/img/fig2a.jpg}
-\includegraphics[width=6.5cm, keepaspectratio,]{res/img/fig2b.jpg}
-\caption{(left) Plot of time (x axis) to find target hotspot location for G (‘o’ symbol) vs.
-NG  (‘x’ symbol) conditions, with time-out declared after 120 sec.
-of unsuccessful search.
-Results are shown for each participant (indicated on the y axis and also using color) and at all 10 target locations.
-(right) Histogram plot showing the distribution of NG time minus G time.
-The plot aggregates time differences for all participants over all locations; positive times mean the search was faster with G than with NG.}
-\label{fig:res}
+\includegraphics[width=12cm, keepaspectratio,]{res/img/hist.png}
+\caption{
+Histogram of the times to find the hotspots for each guidance condition, aggregated over participants and hotspots. The vertical dashed lines represent the median scores.}
+\label{fig:res-ul}
 \end{figure}
 
-We analyzed the differences between times under G vs.~NG using a
-Wilcoxon signed rank test. The resulting p-value is p = 3.8 x 10-8. This
-is a highly statistically significant p-value supporting the rejection
-of the null hypothesis that the distribution of times under G is the
-same as the distribution of times under NG.
+\begin{figure}
+\includegraphics[width=12cm, keepaspectratio,]{res/img/lines.png}
+\caption{
+Line plot showing the average time to find the hotspot ($\pm{SE}$) per subject and per guidance condition.}
+\label{fig:res-ur}
+\end{figure}
 
-The following System Usability Scale (SUS) scores were calculated:
-P1=65, P2=75, P3=77.5, for an average of 72.5. An SUS score of 68 is
-considered ``average'' while scores in the mid- to high 70s are
-considered ``good'' (\url{https://measuringu.com/interpret-sus-score/}).
-Overall, the SUS scores suggest that the guidance system is usable but
-needs improvement.
+\begin{figure}
+\includegraphics[width=12cm, keepaspectratio,]{res/img/full.png}
+\caption{
+Time to find target hotspot location for each participant, colored by guidance condition. This plot displays all the raw experimental data.}
+\label{fig:res-full}
+\end{figure}
 
-Next we summarize the qualitative feedback we obtained in the
-semi-structured interviews. All participants who provided qualitative
-feedback (P1, P2 and P3) acknowledged the benefit of the audio guidance
-that was provided, which they found intuitive and clearly superior to
-finding hotspots without guidance. However, P1 and P2 felt that the
-limited spatial accuracy of the CamIO system was the biggest problem
-that needed to be addressed; for instance, in some cases the system
-announced the target had been reached even though the stylus tip was
-hovering about 1 cm above the truck surface. (This spatial inaccuracy
+These plots show how the time to find the hotspot depends on factors such as the guidance condition (G vs.
+NG), the participant and which hotspot is the target.
+Note that G resulted in a successful target search in all cases, and in all but two cases (both for P4), target search took less time with G than with NG (visible in Fig.
+\ref{fig:res-full}).
+
+We used a Wilcoxon signed-rank test to compare the differences between G and NG times (see Fig.~\ref{fig:res-diff}), paired by hotspot for each participant.
+The result indicated that participants performed significantly better under condition G (median = 14.5 sec.) than in condition NG (median = 120 sec.), with $V=5$, $Z=5.45$, $p = 5.18 \times 10^{-8}$, and an effect size of $r = 0.86$.
+
+\begin{figure}[h!]
+\includegraphics[width=11.5cm, keepaspectratio,]{res/img/hist_diff.png}
+\caption{Histogram of the differences (paired by hotspot \& condition) in the time to find hotspots, aggregated over participants.
+Positive time differences indicate that the time under NG was longer than the time under G.}
+\label{fig:res-diff}
+\end{figure}
+
+It should be noted that, despite our low number of participants, we had 10 pairs of measurements per participant, and a large effect-size (according to Cohen's guidelines), which should theoretically compensate for the increased type-I error rate risk inherent to running inferential statistics on a small sample.
+
+The following System Usability Scale (SUS) scores were calculated: P1=65, P2=75, P3=77.5, for an average of 72.5. 
+An SUS score of 68 is considered ``average'' while scores in the mid- to high 70s are considered ``good''\footnote{\url{https://measuringu.com/interpret-sus-score/}}.
+Overall, the SUS scores suggest that the guidance system is usable but needs improvement.
+
+Next we summarize the qualitative feedback we obtained in the semi-structured interviews. 
+All participants who provided qualitative feedback (P1, P2 and P3) acknowledged the benefit of the audio guidance that was provided, which they found intuitive and clearly superior to finding hotspots without guidance, or to more indirect ways of providing guidance (such as the early version we implemented that used tempo to signal the distance from the target and pitch to direct the stylus up or down).
+P3 specifically favored the ``specific'' instructions that the verbal feedback provided to direct the user to a hotspot, which was especially useful when the hotspot was hard to find using tactile cues alone.
+
+However, P1 and P2 felt that the limited spatial accuracy of the CamIO system was the biggest problem that needed to be addressed; for instance, in some cases the system announced the target had been reached even though the stylus tip was hovering about 1 cm above the truck surface.
+(This spatial inaccuracy
 results from noise in the board pose and stylus tip location estimates.)
-P3 expressed frustration with time lag, i.e., the voice feedback tended
-to lag behind the current location of the stylus, even when she was
-moving the stylus slowly; she also felt that the speed of the repetitive
-speech feedback made her feel rushed.
+P3 expressed frustration with time lag, i.e., the voice feedback tended to lag behind the current location of the stylus, even when she was moving the stylus slowly; she also felt that the speed of the repetitive speech feedback made her feel rushed.
 
-When asked what improvements the participants wanted, P1 suggested
-equipping the stylus with a button that turns on the guidance feedback,
-and the addition of multiple vibrators on the stylus that could provide
-haptic feedback indicating which way the stylus should be moved. P2
-suggested adding audio information similar to the earlier guidance
-system that the authors experimented with (Sec. 3), which would use a
-combination of tempo, pitch and volume changes to signal when the stylus
-gets closer to the target location. P3 expressed the desire that CamIO
-work with any ordinary pen or pencil, rather than requiring the special
-stylus we used.
+When asked what improvements the participants wanted, P1 suggested equipping the stylus with a button that turns on the guidance feedback, and the addition of multiple vibrators on the stylus that could provide haptic feedback indicating which way the stylus should be moved.
+P2 suggested adding audio information similar to the earlier guidance system that the authors experimented with (Sec.
+3), which would use a combination of tempo, pitch and volume changes to signal when the stylus gets closer to the target location.
+P3 expressed the desire that CamIO work with any ordinary pen or pencil, rather than requiring the special stylus we used.
 
@@
 
-We have devised a 3D spatial guidance system that helps a blind user
-find a specific location of interest on an object using CamIO, an
-audio-based AR system. The guidance system uses simple verbal commands
-(``left''/ ``right''/``up''/etc.) to tell the user which direction to
-move the stylus tip. Experiments with four blind participants show that
-the guidance system significantly speeds up the search for target
-hotspots compared with an exhaustive search. A secondary contribution of
-this study is to demonstrate the ability of the CamIO stylus to access
-recessed or otherwise obscured locations on an object that are not
-themselves visible to the camera.
+We have devised a 3D spatial guidance system that helps a blind user find a specific location of interest on an object using CamIO, an audio-based AR system. 
+The guidance system uses simple verbal commands (``left'', ``right'', ``up'', etc.) to tell the user which direction to
+move the stylus tip. Experiments with four blind participants show that the guidance system significantly speeds up the search for target hotspots compared with an exhaustive search. 
+A secondary contribution of this study is to demonstrate the ability of the CamIO stylus to access recessed or otherwise obscured locations on an object that are not themselves visible to the camera.
 
-While we applied this guidance approach specifically to CamIO, it could
-be useful in any AR system, whether the system tracks the user's hands,
-handheld stylus or other pointing tool. Such audio-based guidance is not
-only helpful for blind users but is also useful in applications where
-visibility is limited (either by environmental conditions or because of
-a visual impairment), or for sighted users who prefer a cross-sensory
-display (e.g., audio and visual combined).
+While we applied this guidance approach specifically to CamIO, it could be useful in any AR system, whether the system tracks the user's hands, handheld stylus or other pointing tool. 
+Such audio-based guidance is not only helpful for blind users but is also useful in applications where visibility is limited (either by environmental conditions or because of a visual impairment), or for sighted users who prefer a cross-sensory display (e.g., audio and visual combined).
 
 Future work will focus on refining and improving the audio interface.
-One possible way to speed up the guidance process is to add more
-specific directional feedback, e.g., ``far left'' would tell the user to
-move the stylus farther to the left than the ``left'' instruction. The
-verbal directions could be augmented with 3D spatialized sound as in
-\cite{mayAuditoryDisplaysFacilitate2019}, which could make the guidance
-more intuitive. Finally, we will continue to test and improve the CamIO
-system, with an emphasis on refining the spatial accuracy of the tip
-location estimates using improved computer vision algorithms.
+One possible way to speed up the guidance process is to add more specific directional feedback, e.g., ``far left'' would tell the user to move the stylus farther to the left than the ``left'' instruction.
+The verbal directions could be augmented with 3D spatialized sound as in \cite{mayAuditoryDisplaysFacilitate2019}, which could make the guidance more intuitive. 
+Finally, we will continue to test and improve the CamIO system, with an emphasis on refining the spatial accuracy of the tip location estimates using improved computer vision algorithms.
+
+\section{Acknowledgments}
+JMC, BB and HS were supported by NIH grant 5R01EY025332 and NIDILRR grant 90RE5024-01-00.
+We thank Dr.
+Roberto Manduchi, Dr.
+Natela Shanidze, Dr.
+Santani Teng and Dr.
+Ali Cheraghi for helpful suggestions about the experiments.
 

### 2020-08-04 R/ICCHP20.Rmd
Update renv + utils

@@
 
-if (!require(knitr)) install.packages("knitr")
-if (!require(pacman)) install.packages("pacman")
-pkg_list <- c("tidyverse", "glue", "magrittr", "plotly", "moments", "emmeans", "conflicted", "gridExtra", "kableExtra", "rcompanion", "broom", "rstatix", "ggpubr", "Rmisc")
+source("utils/dependencies.R", echo=F)
 
-pacman::p_load(char = pkg_list) #update=TRUE
+pkg_list <-
+  c(
+    "knitr",
+    "tidyverse",
+    "glue",
+    "magrittr",
+    "plotly",
+    "moments",
+    "emmeans",
+    "conflicted",
+    "gridExtra",
+    "kableExtra",
+    "rcompanion",
+    "broom",
+    "rstatix",
+    "ggpubr",
+    "Rmisc",
+    "lme4"
+  )
+
+update_packages(pkg_list)
+
+options(install.packages.check.source = "no", scipen=999, digits=4)
 
 knitr::opts_chunk$set(
-  warning = FALSE,
+  warning = TRUE,
   echo = FALSE,
-  message = FALSE,
-  error=FALSE,
+  message = TRUE,
+  error = TRUE,
   cache = FALSE
 )
-
-options(scipen=999, digits=4)
 
@@
 ```
+
+## Analysis tests:
+
+```{r}
+lme0 <- lmer(data = data.long, formula = Time ~ Condition + (1|Participant), REML=FALSE)
+lme0 #%>% broom::tidy()
+```
+
+```{r}
+lme1 <- lmer(data = data.long, formula = Time ~ (1|Participant), REML=FALSE)
+lme1
+```
+
+```{r}
+lme2 <- lmer(data = data.long, formula = Time ~ Condition + (Condition|Participant), REML=FALSE)
+lme2
+```
+
+```{r}
+AIC(lme0, lme1, lme2)
+```
+
+```{r}
+anova(lme0, lme1)
+```

### 2020-08-04 R/Index.Rmd
Update renv + utils

@@
 
-if (!require(knitr)) install.packages("knitr")
-if (!require(pacman)) install.packages("pacman")
+source("utils/dependencies.R", echo=F)
+
 pkg_list <- c("tidyverse", "glue", "broom", "magrittr", "prettydoc", "DT", "reshape2", "plotly", "moments", "emmeans", "gtsummary", "lmtest", "table1", "rstatix", "lme4", "ggeffects", "conflicted", "forecast", "Rmisc", "gridExtra", "kableExtra", "rcompanion")
 
-pacman::p_load(char = pkg_list) #update=TRUE
-#pacman::p_load_gh(c("rstudio/gt", "tidyverse/dplyr"))
-#knitr::write_bib(c(.packages(), pkg_list), 'bib/packages.bib')
+update_packages(pkg_list)
 

### 2020-08-04 utils/dependencies.R

@@ -1,0 +1,57 @@
+project_packages <- c()
+
+"%ni%" <- Negate("%in%")
+
+options(install.packages.check.source = "no")
+
+get_pkg_name <- function(pkg) {
+  pkg_name <- pkg
+  if (grepl("/", pkg, fixed = TRUE)) {
+    pkg_path <- stringr::str_split(pkg, "/")[[1]]
+    pkg_name <- pkg_path[length(pkg_path)]
+  }
+  return(pkg_name)
+}
+
+activate_packages <- function() {
+  for (pkg in project_packages) {
+    activate_package(pkg)
+  }
+}
+
+activate_package <- function(pkg) {
+  pkg_name <- get_pkg_name(pkg)
+  if (pkg_name %in% installed.packages()) {
+    library(pkg_name, character.only = TRUE)
+  }
+}
+
+update_packages <- function(pkgs) {
+  for (pkg in pkgs) {
+    if(pkg %ni% project_packages) {
+      project_packages <<- c(project_packages, pkg)
+    }
+  }
+  
+
+  for (pkg in project_packages) {
+    
+    pkg_name <- get_pkg_name(pkg)
+
+    if(!(pkg_name %in% installed.packages())) {
+      if(grepl("/", pkg, fixed=TRUE)) {
+          remotes::install_github(pkg)
+        } else {
+          install.packages(pkg, character.only = TRUE, type = "binary")
+        }
+      
+    }
+    activate_package(pkg)
+  }
+  renv::snapshot(type="all", prompt=F)
+  #knitr::write_bib(c(.packages(), project_packages), here::here("res/bib", "packages.bib"))
+}
+
+update_packages(c("knitr", "renv", "here", "glue", "styler", "remotes"))
+
+# ----------------------------------------------------------------------------

### 2020-09-10 R/ICCHP20.Rmd
Update dep management

--- 2020-06-17 R/ICCHP20.Rmd
@@
 
-if (!require(knitr)) install.packages("knitr")
-if (!require(pacman)) install.packages("pacman")
-pkg_list <- c("tidyverse", "glue", "magrittr", "plotly", "moments", "emmeans", "conflicted", "gridExtra", "kableExtra", "rcompanion", "broom", "rstatix", "ggpubr", "Rmisc")
+source("../utils/dependencies.R", echo=F)
 
-pacman::p_load(char = pkg_list) #update=TRUE
+pkg_list <-
+  c("knitr",
+    "tidyverse",
+    "dplyr",
+    "glue",
+    "magrittr",
+    "plotly",
+    "moments",
+    "emmeans",
+    "conflicted",
+    "gridExtra",
+    "kableExtra",
+    "rcompanion",
+    "broom",
+    "broom.mixed",
+    "afex",
+    "ggpubr",
+    "Rmisc",
+    "lme4",
+    "survival",
+    "coxme",
+    "survminer",
+    "easystats/see",
+    "easystats/report",
+    "easystats/performance",
+    "stargazer",
+    "jtools",
+    "gtools"
+  )
+
+update_packages(pkg_list)
+
+options(install.packages.check.source = "no", scipen=999, digits=4)
 
 knitr::opts_chunk$set(
-  warning = FALSE,
+  warning = TRUE,
   echo = FALSE,
-  message = FALSE,
-  error=FALSE,
+  message = TRUE,
+  error = TRUE,
   cache = FALSE
@@
 
-options(scipen=999, digits=4)
-
 alpha <- 0.05
 set.seed(42)
-
-initial.data <- "../data/initial.csv"
 
@@
 conflict_prefer("layout", "plotly")
+conflict_prefer("here", "here")
+
+initial.data <- here("data", "initial.csv")
 
@@
   rename("Condition" = mode, "Hotspot" = hotspot) %>%
-  mutate(across(c(Hotspot, Condition), as.factor)) %T>% print
+  mutate(across(c(Hotspot, Condition), as.factor)) #%T>% print
 
@@
   pivot_longer(cols = "P1":"P4", names_to="Participant", names_transform = list(Participant = as.factor), values_to = "Time") %>% 
-  relocate(Participant, id, Condition, Hotspot, Time) %>%
+  mutate(Success = as.integer(ifelse(Time >= 120, 0, 1))) %>%
+  relocate(Participant, id, Condition, Hotspot, Success, Time) %>%
   #mutate(Hotspot = factor(Hotspot, levels = hs.order)) %>%
@@
 ```
+
 ***
@@
 ```
+
+<!--------------------------------------------------------------------------------->
+<!--------------------------------------------------------------------------------->
+
+## Analysis tests:
+***
+
+### Model building:
+
+```{r}
+cox.mod <- coxph(Surv(Time, Success) ~ Condition + frailty.gaussian(Participant) + frailty.gaussian(Hotspot), data = data.long)
+cox.mod.min <- coxph(Surv(Time, Success) ~ 1 + frailty.gaussian(Participant) + frailty.gaussian(Hotspot), data = data.long)
+summary(cox.mod)
+```
+
+```{r}
+model_performance(cox.mod)
+```
+
+```{r}
+anova(cox.mod, cox.mod.min)
+```
+
+### With cox.me:
+
+```{r}
+cox_me <- coxme(Surv(Time, Success) ~ Condition + (1 | Participant) + (1 | Hotspot), data = data.long)
+cox_me.min <- coxme(Surv(Time, Success) ~ 1 + (1 | Participant) + (1 | Hotspot), data = data.long)
+summary(cox.me)
+```
+
+```{r}
+AIC(cox.me)
+```
+
+```{r}
+anova(cox_me, cox_me.min)
+```
+
+### Model evaluation:
+
+```{r}
+# ggcoxzph(cox.zph(cox.mod))
+```
+
+```{r}
+ggcoxdiagnostics(cox.mod,
+ type = "schoenfeld",
+ ox.scale = "linear.predictions")
+```
+
+```{r}
+ggforest(cox.mod)
+```

### 2020-09-10 utils/dependencies.R

@@ -1,8 +1,42 @@
-project_packages <- c()
+#########################
+# Managing dependencies #
+#########################
+
+set.github_pat <- function(env.var) {
+  github.pat <- Sys.getenv(env.var)
+  if (github.pat != "") {
+    print(paste("[INFO] Found GITHUB Access Token: ", github.pat))
+    GITHUB_PAT <- github.pat
+  }
+}
+
+set.install_type <- function() {
+  sys.name <- Sys.info()[["sysname"]]
+  if (sys.name == "Windows") {
+    print("[INFO] Windows detected, installing packages as 'binary'")
+    options(install.packages.check.source = "no")
+    return("binary")
+  } else if (sys.name == "Linux") {
+    print("[INFO] Linux detected, installing packages as 'source'")
+    options(install.packages.check.source = "yes")
+    return("source")
+  } else {
+    print("[INFO] No install type setup for your system, using 'source' as default")
+    options(install.packages.check.source = "yes")
+    return("source")
+  }
+}
+
+set.github_pat("GITHUB_PAT_R_INSTALL")
+pkg.install.type <- set.install_type()
 
 "%ni%" <- Negate("%in%")
 
-options(install.packages.check.source = "no")
+Sys.setenv(MAKEFLAGS = "-j4")
+
+# -------------------------------------------
+
+project_packages <- c()
 
 get_pkg_name <- function(pkg) {
   pkg_name <- pkg
@@ -22,7 +56,7 @@ activate_packages <- function() {
 activate_package <- function(pkg) {
   pkg_name <- get_pkg_name(pkg)
   if (pkg_name %in% installed.packages()) {
-    library(pkg_name, character.only = TRUE)
+    library(pkg_name, character.only = TRUE, quiet=TRUE)
   }
 }
 
@@ -33,17 +67,17 @@ update_packages <- function(pkgs) {
     }
   }
   
-
+  
   for (pkg in project_packages) {
     
     pkg_name <- get_pkg_name(pkg)
-
+    
     if(!(pkg_name %in% installed.packages())) {
       if(grepl("/", pkg, fixed=TRUE)) {
-          remotes::install_github(pkg)
-        } else {
-          install.packages(pkg, character.only = TRUE, type = "binary")
-        }
+        remotes::install_github(pkg, upgrade = "never", quiet = TRUE)
+      } else {
+        install.packages(pkg, character.only = TRUE, type = pkg.install.type, quiet = TRUE, verbose = FALSE)
+      }
[…]
